---
title: "Hands-on_Ex0402"
author: "JUNCHEN LI"
date: "May 7,2026"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: auto
---

## 10.3.1 Installing and launching R packages

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

## 10.3.2 Importing data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## 10.3.3 One-sample test: gghistostats() method

::: panel-tabset
## Ex plot

```{r, echo=FALSE}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

## The Code

```{r, eval=FALSE}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```
:::

## Self Trying

::: Panel-tabset
## The plot

```{r, echo=FALSE}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = SCIENCE,
  type = "bayes",
  test.value = 60,
  xlab = "SCIENCE scores"
)
```

## The Code

```{r, eval=FALSE}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = SCIENCE,
  type = "bayes",
  test.value = 60,
  xlab = "SCIENCE scores"
)
```
:::

## 10.3.6 Two-sample mean test: ggbetweenstats()

::: Panel-tabset
## The Ex plot

```{r, echo=FALSE}
ggbetweenstats(
  data = exam_data,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

## The Code

```{r, eval=FALSE}
ggbetweenstats(
  data = exam_data,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```
:::

## Self Trying

::: Panel-tabset
## The Plot

```{r, echo=FALSE}
ggbetweenstats(
  data = exam_data,
  x = RACE, 
  y = ENGLISH,
  type = "np",
  messages = FALSE
)
```

## The Code

```{r, eval=FALSE}
ggbetweenstats(
  data = exam_data,
  x = RACE, 
  y = ENGLISH,
  type = "np",
  messages = FALSE
)
```
:::

## 10.3.7 Oneway ANOVA Test: ggbetweenstats() method

::: panel-tabset
## The Ex plot

```{r, echo=FALSE}
ggbetweenstats(
  data = exam_data,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

## The Code

```{r, eval=FALSE}
ggbetweenstats(
  data = exam_data,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```
:::

## Self Trying

::: panel-tabset
## The plot

```{r, echo=FALSE}
ggbetweenstats(
  data = exam_data,
  x = GENDER, 
  y = SCIENCE,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

## The code

```{r, eval=FALSE}
ggbetweenstats(
  data = exam_data,
  x = GENDER, 
  y = SCIENCE,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```
:::

## 10.3.8 Significant Test of Correlation: ggscatterstats()

::: panel-tabset
## The plot

```{r, echo=FALSE}
ggscatterstats(
  data = exam_data,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

## The Code

```{r, eval=FALSE}
ggscatterstats(
  data = exam_data,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```
:::

## Self Trying

::: panel-tabset
## The plot

```{r, echo=FALSE}
ggscatterstats(
  data = exam_data,
  x = SCIENCE,
  y = MATHS,
  marginal = FALSE,
  )
```

## The Code

```{r, eval=FALSE}
ggscatterstats(
  data = exam_data,
  x = SCIENCE,
  y = MATHS,
  marginal = FALSE,
  )
```
:::

## 10.3.9 Significant Test of Association (Depedence) : ggbarstats() methods

::: panel-tabset
## The Ex plot

```{r, echo=FALSE}
exam1 <- exam_data %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```

## The Code

```{r, eval=FALSE}
exam1 <- exam_data %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```
:::

## Self-Trying

::: panel-tabset
## The plot

```{r, echo=FALSE}
exam1 <- exam_data %>% 
  mutate(SCIENCE_bins = 
           cut(SCIENCE, 
               breaks = c(0,60,75,85,100))
)
ggbarstats(exam1, 
           x = SCIENCE_bins, 
           y = RACE)
```

## The Code

```{r, eval=FALSE}
exam1 <- exam_data %>% 
  mutate(SCIENCE_bins = 
           cut(SCIENCE, 
               breaks = c(0,60,75,85,100))
)
ggbarstats(exam1, 
           x = SCIENCE_bins, 
           y = RACE)
```
:::
