---
title: "Hands-on Exercise: Visualising and Analysing Time-oriented Data"
author: "JUNCHEN LI"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    df-print: paged
page-layout: full
execute:
  warning: false
  message: false
---

# Overview

This hands-on exercise follows the learning flow of Chapter 17 of *R for Visual Analytics*. The teacher example introduces three approaches for visualising time-oriented data: calendar heatmaps, cycle plots, and slopegraphs. The Self Trying section extends these methods with additional comparisons and a clearer analytical focus.

The project folder already includes the three data files required for this exercise:

``` text
data/eventlog.csv
data/arrivals_by_air.xlsx
data/rice.csv
```

The uploaded datasets were checked before packaging. The event log contains 199,999 records, the visitor-arrival workbook covers January 2000 to December 2019, and the rice dataset covers 11 countries from 1961 to 2010.

# 1 Getting Started

## 1.1 Loading R packages

```{r}
packages <- c(
  "scales", "viridis", "lubridate", "ggthemes",
  "gridExtra", "readxl", "knitr", "data.table",
  "tidyverse", "DT", "ggrepel"
)

pacman::p_load(char = packages)
```

# Teacher Example

# 2 Plotting Calendar Heatmaps

## 2.1 Importing the event log data

The event log records cyber attacks. Each row contains a timestamp, a source country, and the time zone associated with the source IP address.

```{r}
attacks_raw <- read_csv("data/eventlog.csv", show_col_types = FALSE)
```

## 2.2 Examining the data structure

```{r}
kable(head(attacks_raw))
```

## 2.3 Preparing weekday and hour fields

To avoid errors caused by different Windows language settings, the code derives weekday labels from numeric weekday values rather than relying on the computer locale.

```{r}
make_hr_wkday <- function(timestamp, source_country, tz) {
  
  local_time <- lubridate::with_tz(
    timestamp,
    tzone = tz[1]
  )
  
  tibble(
    source_country = source_country,
    wkday_num = ((lubridate::wday(local_time) + 5) %% 7) + 1,
    hour = lubridate::hour(local_time)
  )
}

wkday_labels <- c(
  "Monday", "Tuesday", "Wednesday", "Thursday",
  "Friday", "Saturday", "Sunday"
)

wkday_levels <- c(
  "Saturday", "Friday", "Thursday", "Wednesday",
  "Tuesday", "Monday", "Sunday"
)

attacks <- attacks_raw %>%
  group_by(tz) %>%
  summarise(
    processed = list(
      make_hr_wkday(timestamp, source_country, tz)
    ),
    .groups = "drop"
  ) %>%
  unnest(processed) %>%
  mutate(
    wkday = factor(
      wkday_labels[wkday_num],
      levels = wkday_levels
    ),
    hour = factor(hour, levels = 0:23)
  )
```

```{r}
kable(head(attacks))
```

## 2.4 Building a basic calendar heatmap

The first heatmap shows the overall distribution of attacks by weekday and hour of day.

```{r}
grouped_attacks <- attacks %>%
  count(wkday, hour) %>%
  drop_na()

ggplot(
  grouped_attacks,
  aes(x = hour, y = wkday, fill = n)
) +
  geom_tile(colour = "white", linewidth = 0.1) +
  coord_equal() +
  scale_fill_gradient(
    name = "Number of attacks",
    low = "skyblue",
    high = "darkblue"
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Attacks by Weekday and Time of Day"
  ) +
  theme_tufte(base_family = "Arial") +
  theme(
    axis.ticks = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 7)
  )
```

## 2.5 Identifying the top four source countries

```{r}
attacks_by_country <- attacks %>%
  count(source_country, sort = TRUE) %>%
  mutate(percent = percent(n / sum(n)))

kable(head(attacks_by_country, 10))
```

## 2.6 Preparing the top-four-country data

```{r}
top4 <- attacks_by_country %>%
  slice_head(n = 4) %>%
  pull(source_country)

top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  mutate(
    source_country = factor(
      source_country,
      levels = top4
    )
  ) %>%
  drop_na()
```

## 2.7 Plotting multiple calendar heatmaps

```{r}
ggplot(
  top4_attacks,
  aes(x = hour, y = wkday, fill = n)
) +
  geom_tile(colour = "white", linewidth = 0.1) +
  coord_equal() +
  scale_fill_gradient(
    name = "Number of attacks",
    low = "skyblue",
    high = "darkblue"
  ) +
  facet_wrap(~source_country, ncol = 2) +
  labs(
    x = NULL,
    y = NULL,
    title = "Attacks from the Top Four Source Countries by Weekday and Hour"
  ) +
  theme_tufte(base_family = "Arial") +
  theme(
    axis.ticks = element_blank(),
    axis.text.x = element_text(size = 7),
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 7)
  )
```

# 3 Plotting a Cycle Plot

## 3.1 Importing visitor-arrival data

```{r}
air <- read_excel("data/arrivals_by_air.xlsx") %>%
  mutate(
    `Month-Year` = as.Date(`Month-Year`),
    month = factor(
      month(`Month-Year`),
      levels = 1:12,
      labels = month.abb,
      ordered = TRUE
    ),
    year = year(`Month-Year`)
  )
```

## 3.2 Extracting Vietnam arrivals

```{r}
Vietnam <- air %>%
  transmute(
    arrivals = as.numeric(.data[["Vietnam"]]),
    month,
    year
  ) %>%
  filter(year >= 2010)
```

## 3.3 Computing monthly averages

```{r}
vietnam_monthly_average <- Vietnam %>%
  group_by(month) %>%
  summarise(
    avgvalue = mean(arrivals, na.rm = TRUE),
    .groups = "drop"
  )
```

## 3.4 Plotting the cycle plot

```{r}
ggplot() +
  geom_line(
    data = Vietnam,
    aes(
      x = year,
      y = arrivals,
      group = month
    ),
    colour = "black"
  ) +
  geom_hline(
    data = vietnam_monthly_average,
    aes(yintercept = avgvalue),
    linetype = 6,
    colour = "red",
    linewidth = 0.5
  ) +
  facet_grid(~month) +
  labs(
    x = NULL,
    y = "Number of Visitors",
    title = "Visitor Arrivals from Vietnam by Air"
  ) +
  theme_tufte(base_family = "Arial") +
  theme(axis.text.x = element_blank())
```

# 4 Plotting a Slopegraph

## 4.1 Importing rice-yield data

```{r}
rice <- read_csv("data/rice.csv", show_col_types = FALSE)
```

## 4.2 Preparing the comparison

The slopegraph compares rice yield in 1961 and 1980. A direct `ggplot2` implementation is used so that the file remains easy to render without additional specialised packages.

```{r}
rice_teacher <- rice %>%
  filter(Year %in% c(1961, 1980)) %>%
  mutate(Year = factor(Year))
```

## 4.3 Plotting the slopegraph

```{r}
ggplot(
  rice_teacher,
  aes(
    x = Year,
    y = Yield,
    group = Country
  )
) +
  geom_line(colour = "grey60", linewidth = 0.7) +
  geom_point(size = 2.5) +
  geom_text_repel(
    aes(label = Country),
    size = 3,
    direction = "y",
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  labs(
    x = NULL,
    y = "Rice Yield",
    title = "Rice Yield of Asian Countries",
    subtitle = "Comparison between 1961 and 1980"
  ) +
  theme_minimal(base_family = "Arial")
```

# Self Trying

In this section, I extend the teacher example with additional analysis. Instead of repeating the same charts, I focus on comparability and storytelling. First, I compare attack timing patterns after standardising the attack volume within each country. Next, I identify peak attack periods. I then apply the cycle-plot concept to another visitor market and use a longer time horizon to analyse rice-yield changes.

# 5 Self Trying: Comparing Attack Timing Patterns

## 5.1 Selecting the top six source countries

The teacher example compares the top four countries using raw attack counts. My analysis expands the comparison to six countries and calculates within-country percentages. This prevents countries with larger attack volumes from dominating the visual interpretation.

```{r}
top6 <- attacks_by_country %>%
  slice_head(n = 6) %>%
  pull(source_country)

top6_profile <- attacks %>%
  filter(source_country %in% top6) %>%
  count(source_country, wkday, hour) %>%
  group_by(source_country) %>%
  mutate(
    country_total = sum(n),
    share = n / country_total
  ) %>%
  ungroup() %>%
  mutate(
    source_country = factor(
      source_country,
      levels = top6
    )
  ) %>%
  drop_na()
```

## 5.2 Plotting normalised calendar heatmaps

```{r}
ggplot(
  top6_profile,
  aes(x = hour, y = wkday, fill = share)
) +
  geom_tile(colour = "white", linewidth = 0.1) +
  coord_equal() +
  scale_fill_viridis_c(
    labels = percent_format(accuracy = 0.1),
    name = "Share of each\ncountry's attacks"
  ) +
  facet_wrap(~source_country, ncol = 3) +
  labs(
    x = NULL,
    y = NULL,
    title = "Normalised Attack Timing Profiles for the Top Six Source Countries",
    subtitle = "Each heatmap shows the within-country percentage distribution of attacks"
  ) +
  theme_tufte(base_family = "Arial") +
  theme(
    axis.ticks = element_blank(),
    axis.text.x = element_text(size = 6),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5)
  )
```

## 5.3 Identifying peak attack periods

```{r}
peak_attack_periods <- top6_profile %>%
  group_by(source_country) %>%
  slice_max(
    order_by = n,
    n = 1,
    with_ties = FALSE
  ) %>%
  ungroup() %>%
  transmute(
    source_country,
    peak_weekday = wkday,
    peak_hour = as.character(hour),
    attacks = n,
    within_country_share = percent(share, accuracy = 0.1)
  )

DT::datatable(
  peak_attack_periods,
  rownames = FALSE,
  options = list(
    pageLength = 6,
    dom = "tip"
  )
)
```

## 5.4 Comparing the busiest hours

```{r}
peak_attack_periods %>%
  mutate(
    peak_hour = as.numeric(peak_hour),
    source_country = fct_reorder(source_country, peak_hour)
  ) %>%
  ggplot(
    aes(
      x = peak_hour,
      y = source_country
    )
  ) +
  geom_segment(
    aes(
      x = 0,
      xend = peak_hour,
      yend = source_country
    ),
    linewidth = 0.8,
    colour = "grey70"
  ) +
  geom_point(size = 4) +
  scale_x_continuous(breaks = 0:23) +
  labs(
    x = "Peak Hour of Day",
    y = NULL,
    title = "Peak Attack Hour for Each Leading Source Country"
  ) +
  theme_minimal(base_family = "Arial")
```

## 5.5 Interpretation

The normalised heatmaps support fairer comparisons across countries because the colour scale represents each country's internal attack distribution rather than its absolute number of attacks. The peak-period table and lollipop chart make the most active time window easier to identify.

# 6 Self Trying: Exploring Visitor Seasonality for Another Market

## 6.1 Selecting a second visitor market

The code uses China when the column exists. If the imported workbook does not contain a China column, it automatically selects the first available country column.

```{r}
available_markets <- setdiff(
  names(air),
  c("Month-Year", "month", "year")
)

target_country <- if ("China" %in% available_markets) {
  "China"
} else {
  available_markets[1]
}

target_country
```

## 6.2 Preparing the target-market data

```{r}
selected_market <- air %>%
  transmute(
    arrivals = as.numeric(.data[[target_country]]),
    month,
    year
  ) %>%
  filter(year >= 2010)

selected_market_average <- selected_market %>%
  group_by(month) %>%
  summarise(
    avgvalue = mean(arrivals, na.rm = TRUE),
    .groups = "drop"
  )
```

## 6.3 Plotting the cycle plot

```{r}
ggplot() +
  geom_line(
    data = selected_market,
    aes(
      x = year,
      y = arrivals,
      group = month
    ),
    colour = "grey30"
  ) +
  geom_hline(
    data = selected_market_average,
    aes(yintercept = avgvalue),
    linetype = 6,
    colour = "red",
    linewidth = 0.5
  ) +
  facet_grid(~month) +
  labs(
    x = NULL,
    y = "Number of Visitors",
    title = paste("Visitor Arrivals from", target_country, "by Air"),
    subtitle = "The dashed line shows the average arrival level for each month"
  ) +
  theme_tufte(base_family = "Arial") +
  theme(axis.text.x = element_blank())
```

## 6.4 Plotting the average monthly profile

```{r}
selected_market_average %>%
  ggplot(
    aes(
      x = month,
      y = avgvalue,
      group = 1
    )
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  labs(
    x = NULL,
    y = "Average Number of Visitors",
    title = paste("Average Monthly Arrival Pattern for", target_country)
  ) +
  theme_minimal(base_family = "Arial")
```

## 6.5 Interpretation

The second cycle plot applies the same method to another visitor market. The monthly-profile line chart complements the cycle plot by making seasonal peaks and low seasons easier to compare.

# 7 Self Trying: Analysing Long-term Rice Yield Changes

## 7.1 Selecting the earliest and latest years

Instead of using a fixed pair of years, I compare the earliest and latest years available in the rice dataset. This makes the analysis reusable if the dataset is updated.

```{r}
first_year <- min(rice$Year, na.rm = TRUE)
last_year <- max(rice$Year, na.rm = TRUE)

c(first_year, last_year)
```

## 7.2 Computing yield changes

```{r}
rice_change <- rice %>%
  filter(Year %in% c(first_year, last_year)) %>%
  group_by(Country) %>%
  summarise(
    start_yield = Yield[Year == first_year][1],
    end_yield = Yield[Year == last_year][1],
    yield_change = end_yield - start_yield,
    .groups = "drop"
  ) %>%
  drop_na()

top_change_countries <- rice_change %>%
  slice_max(
    order_by = abs(yield_change),
    n = 10,
    with_ties = FALSE
  )
```

## 7.3 Plotting the yield-change bar chart

```{r}
top_change_countries %>%
  mutate(
    Country = fct_reorder(Country, yield_change)
  ) %>%
  ggplot(
    aes(
      x = yield_change,
      y = Country,
      fill = yield_change > 0
    )
  ) +
  geom_col(show.legend = FALSE) +
  geom_vline(xintercept = 0, colour = "grey40") +
  labs(
    x = paste("Yield Change:", first_year, "to", last_year),
    y = NULL,
    title = "Countries with the Largest Long-term Rice Yield Changes"
  ) +
  theme_minimal(base_family = "Arial")
```

## 7.4 Plotting a focused slopegraph

```{r}
rice_long_term <- rice %>%
  filter(
    Country %in% top_change_countries$Country,
    Year %in% c(first_year, last_year)
  ) %>%
  mutate(Year = factor(Year))

ggplot(
  rice_long_term,
  aes(
    x = Year,
    y = Yield,
    group = Country
  )
) +
  geom_line(colour = "grey65", linewidth = 0.8) +
  geom_point(size = 2.5) +
  geom_text_repel(
    aes(label = Country),
    direction = "y",
    size = 3,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  labs(
    x = NULL,
    y = "Rice Yield",
    title = "Long-term Rice Yield Change for Selected Countries",
    subtitle = paste(first_year, "to", last_year)
  ) +
  theme_minimal(base_family = "Arial")
```

## 7.5 Displaying the comparison table

```{r}
DT::datatable(
  top_change_countries %>%
    arrange(desc(abs(yield_change))),
  rownames = FALSE,
  options = list(
    pageLength = 10,
    dom = "tip"
  )
)
```

## 7.6 Interpretation

The bar chart quickly identifies the countries with the largest long-term increases or decreases in yield. The slopegraph then shows the direction and magnitude of change in a compact visual form. Using the earliest and latest available years provides a broader view than the fixed two-year teacher example.

# 8 Summary of the Self Trying Approach

The Self Trying section extends the teacher example in three ways:

1.  It uses within-country percentages to compare attack timing patterns fairly across countries with different attack volumes.
2.  It complements the calendar heatmap with a peak-period table and lollipop chart.
3.  It applies the cycle-plot concept to another visitor market and expands the slopegraph analysis to the full available rice-yield period.

These additions preserve the original learning objectives while providing a clearer and more independent analytical narrative.
