---
title: "Hands-on_Ex0504"
author: "JUNCHEN LI"
date: "May 14,2026"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: auto
---

## 15.2 Installing and Launching R Packages

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

## 15.3 Data Preparation

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

## 15.4.1 Plotting a simple parallel coordinates

```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

### Self Trying: Parallel coordinates plot for top 30 happiest countries

```{r}
wh_parallel <- wh %>%
  mutate(
    happiness_group = ifelse(
      `Happiness score` >= median(`Happiness score`, na.rm = TRUE),
      "Higher happiness",
      "Lower happiness"
    )
  ) %>%
  select(
    Country,
    happiness_group,
    GDP_per_capita = `GDP per capita`,
    Social_support = `Social support`,
    Healthy_life_expectancy = `Healthy life expectancy`,
    Freedom = `Freedom to make life choices`,
    Generosity,
    Corruption = `Perceptions of corruption`
  )

ggparcoord(
  data = wh_parallel,
  columns = 3:8,
  groupColumn = 2,
  scale = "uniminmax",
  alphaLines = 0.7
)
```

## 15.4.2 Plotting a parallel coordinates with boxplot

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```

### Self Trying

```{r}
ggparcoord(
  data = wh_parallel,
  columns = 2:7,
  groupColumn = 1,
  scale = "uniminmax",
  alphaLines = 0.2,
  boxplot = TRUE,
  title = "Parallel Coordinates Plot by Happiness Group"
)
```

## 15.4.3 Parallel coordinates with facet

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)
```

### Self Trying: Parallel coordinates with facet by happiness group

```{r}
ggparcoord(
  data = wh_parallel,
  columns = 3:8,
  groupColumn = 2,
  scale = "uniminmax",
  alphaLines = 0.2,
  boxplot = TRUE,
  title = "Faceted Parallel Coordinates Plot by Happiness Group"
) +
  facet_wrap(~ happiness_group)
```

## 15.4.4 Rotating x-axis text label

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))
```

### Self Trying

```{r}
ggparcoord(data = wh_parallel, 
           columns = 3:8, 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Faceted Parallel Coordinates Plot by Happiness Group") +
  facet_wrap(~ happiness_group) + 
  theme(axis.text.x = element_text(angle = 30))
```

## 15.4.5 Adjusting the rotated x-axis text label

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

### Self Trying

```{r}
ggparcoord(data = wh_parallel, 
           columns = 3:8, 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Faceted Parallel Coordinates Plot by Happiness Group") +
  facet_wrap(~ happiness_group) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

## 15.5.1 The basic plot

```{r}
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)
```

### Self Trying: Interactive parallel coordinates plot for top 30 happiest countries

```{r}
wh_parallel_interactive <- wh %>%
  select(
    Happiness_score = `Happiness score`,
    GDP_per_capita = `GDP per capita`,
    Social_support = `Social support`,
    Healthy_life_expectancy = `Healthy life expectancy`,
    Freedom = `Freedom to make life choices`,
    Generosity,
    Corruption = `Perceptions of corruption`
  )

parallelPlot(
  wh_parallel_interactive,
  width = 500,
  height = 350
)
```

## 15.5.2 Rotate axis label

```{r}
parallelPlot(wh,
             rotateTitle = TRUE)
```

### Self Trying

```{r}
parallelPlot(wh_parallel_interactive,
             rotateTitle = TRUE)
```

## 15.5.3 Changing the colour scheme

```{r}
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

### Self Trying

```{r}
parallelPlot(wh_parallel_interactive,
             continuousCS = "YlGnBu",
             rotateTitle = TRUE)
```

## 15.5.4 Parallel coordinates plot with histogram

```{r}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```

### Self Trying

```{r}
histoVisibility <- rep(TRUE, ncol(wh_parallel_interactive))
parallelPlot(wh_parallel_interactive,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```
