---
title: "Hands-on_Ex0502"
author: "JUNCHEN LI"
date: "May 14,2026"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: auto
---

## 6.2 Installing and Launching R Packages

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)
```

## 6.3.1 Importing Data

```{r}
wine <- read_csv("data/wine_quality.csv")
```

## 6.4.1 Building a basic correlation matrix

```{r}
pairs(wine[,1:11])
```

### Scatterplot matrix

```{r}
pairs(wine[,2:12])
```

## 6.4.2 Drawing the lower corner

```{r}
pairs(wine[,2:12], upper.panel = NULL)
```

```{r}
pairs(wine[,2:12], lower.panel = NULL)
```

## 6.4.3 Including with correlation coefficients

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)
```

### Self Tyring

```{r}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...) {
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  
  r <- abs(cor(x, y, use = "complete.obs"))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste(prefix, txt, sep = "")
  
  if (missing(cex.cor)) {
    cex.cor <- 0.8 / strwidth(txt)
  }
  
  text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

wine_selected <- wine[, c(
  "fixed acidity",
  "volatile acidity",
  "citric acid",
  "alcohol",
  "sulphates",
  "quality"
)]

pairs(
  wine_selected,
  upper.panel = panel.cor,
  main = "Correlation Matrix for Selected Wine Quality Variables"
)
```

## 6.5.1 The basic plot

```{r}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11)
```

```{r}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  title    = "Correlogram for wine dataset",
  subtitle = "Four pairs are no significant at p < 0.05"
)
```

```{r}
ggplot.component = list(
    theme(text=element_text(size=5),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8)))
```

### Self Trying: Correlation plot for selected wine variables

```{r}
ggstatsplot::ggcorrmat(
  data = wine,
  cor.vars = c(
    "fixed acidity",
    "volatile acidity",
    "citric acid",
    "residual sugar",
    "alcohol",
    "quality"
  ),
  ggcorrplot.args = list(
    outline.color = "black",
    hc.order = TRUE,
    tl.cex = 10
  ),
  title = "Correlogram for Selected Wine Quality Variables",
  subtitle = "Selected chemical variables and wine quality"
) +
  theme(
    text = element_text(size = 6),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8)
  )
```

## 6.6 Building multiple plots

```{r}
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

### Self Trying

```{r}
## Self Trying: Grouped correlation plot for selected wine variables

grouped_ggcorrmat(
  data = wine,
  cor.vars = c(
    `fixed acidity`,
    `volatile acidity`,
    `citric acid`,
    `residual sugar`,
    alcohol,
    quality
  ),
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(
    outline.color = "grey40",
    hc.order = TRUE,
    tl.cex = 10),
  annotation.args = list(
    tag_levels = "A",
    title = "Grouped Correlation Matrix for Selected Wine Variables",
    subtitle = "Comparing correlation patterns by wine type",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

## 6.7.1 Getting started with corrplot

```{r}
wine.cor <- cor(wine[, 1:11])
```

```{r}
corrplot(wine.cor)
```

## 6.7.2 Working with visual geometrics

```{r}
corrplot(wine.cor, 
         method = "ellipse") 
```

### Self Trying

```{r}
corrplot(wine.cor,
         method = "square")
```

## 6.7.3 Working with layout

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower")
```

### Self Tyring

```{r}
corrplot(wine.cor, 
         method = "square", 
         type="upper")
```

### Ex plot

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower",
         diag = FALSE,
         tl.col = "black")
```

### Self Trying plot

```{r}
corrplot(wine.cor, 
         method = "square", 
         type="upper",
         diag = FALSE,
         tl.col = "red")
```

## 6.7.4 Working with mixed layout

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

### Self Trying

```{r}
corrplot.mixed(wine.cor, 
               lower = "pie", 
               upper = "color",
               tl.pos = "lt",
               diag = "l",
               tl.col = "red")
```

## 6.7.5 Combining corrgram with the significant test

```{r}
wine.sig = cor.mtest(wine.cor, conf.level= .95)
```

```{r}
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)
```

### Self Trying

```{r}
corrplot(wine.cor,
         method = "pie",
         type = "upper",
         diag = FALSE,
         tl.col = "red",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)
```

## 6.7.6 Reorder a corrgram

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "black")
```

### Self Trying

```{r}
corrplot.mixed(wine.cor, 
               lower = "square", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "red")
```

## 6.7.7 Reordering a correlation matrix using hclust

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```

### Self Trying

```{r}
corrplot(wine.cor, 
         method = "square", 
         tl.pos = "lt",
         tl.col = "red",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```
