---
title: "Hands-on_Ex0503"
author: "JUNCHEN LI"
date: "May 14,2026"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: auto
---

## 14.2 Installing and Launching R Packages

```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
```

## 14.3.1 Importing the data set

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

## 14.3.2 Preparing the data

```{r}
row.names(wh) <- wh$Country
```

## 14.3.3 Transforming the data frame into a matrix

```{r}
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

## 14.4.1 heatmap() of R Stats

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)
```

```{r}
wh_heatmap <- heatmap(wh_matrix)
```

### Self Trying

#### 14.3.2 Preparing the data

```{r}
library(dplyr)

wh_top30 <- wh %>%
  arrange(desc(`Happiness score`)) %>%
  slice(1:30)

row.names(wh_top30) <- wh_top30$Country
```

#### 14.3.3 Transforming the data frame into a matrix

```{r}
wh_top30_selected <- wh_top30 %>%
  dplyr::select(
    `Happiness score`,
    `GDP per capita`,
    `Social support`,
    `Healthy life expectancy`,
    `Freedom to make life choices`,
    `Generosity`,
    `Perceptions of corruption`
  )

wh_top30_matrix <- data.matrix(wh_top30_selected)
```

#### 14.4.1 heatmap() of R Stats

```{r}
wh_top30_heatmap <- heatmap(
  wh_top30_matrix,
  Rowv = NA,
  Colv = NA,
  scale = "column"
)
```

```{r}
wh_top30_heatmap <- heatmap(wh_top30_matrix)
```

## Ex plot

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

### Self Trying

```{r}
wh_top30_heatmap <- heatmap(wh_top30_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

## 14.5.1 Working with heatmaply

```{r}
heatmaply(mtcars)
```

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```

### Self Trying

```{r}
heatmaply(wh_top30_matrix[, -c(1, 2, 4, 5)])
```

## 14.5.2 Data trasformation

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))
```

```{r}
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```

### Self Trying

```{r}
heatmaply(wh_top30_matrix[, -c(1, 2, 4, 5)],
          scale = "column")
```

```{r}
heatmaply(normalize(wh_top30_matrix[, -c(1, 2, 4, 5)]))
```

```{r}
heatmaply(percentize(wh_top30_matrix[, -c(1, 2, 4, 5)]))
```

## 14.5.4 Manual approach

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```

### Self Trying

```{r}
heatmaply(normalize(wh_top30_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "manhattan",
          hclust_method = "ward.D")
```

## 14.5.5 Statistical approach

```{r}
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]
```

```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```

### Self Trying

```{r}
wh_d <- dist(normalize(wh_top30_matrix[, -c(1, 2, 4, 5)]), method = "manhattan")
dend_expend(wh_d)[[3]]
```

```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

```{r}
heatmaply(normalize(wh_top30_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "manhattan",
          hclust_method = "average",
          k_row = 3)
```

## 14.5.6 Seriation

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "mean")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none")
```

### Self Trying

```{r}
heatmaply(normalize(wh_top30_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```

```{r}
heatmaply(normalize(wh_top30_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```

```{r}
heatmaply(normalize(wh_top30_matrix[, -c(1, 2, 4, 5)]),
          seriate = "mean")
```

```{r}
heatmaply(normalize(wh_top30_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none")
```

## 14.5.7 Working with colour palettes

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues)
```

### Self Trying

```{r}
heatmaply(normalize(wh_top30_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Reds)
```

## 14.5.8 The finishing touch

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```

### Self Trying

```{r}
heatmaply(normalize(wh_top30_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Reds,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="Top 30 World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```
