---
title: "Hands-on_Ex0505"
author: "JUNCHEN LI"
date: "May 14,2026"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: auto
---

## 16.2 Installing and Launching R Packages

```{r}
pacman::p_load(treemap, treemapify, tidyverse) 
```

## 16.3.1 Importing the data set

```{r}
realis2018 <- read_csv("data/realis2018.csv")
```

## 16.3.3 Grouped summaries without the Pipe

```{r}
realis2018_grouped <- group_by(realis2018, `Project Name`,
                               `Planning Region`, `Planning Area`, 
                               `Property Type`, `Type of Sale`)
realis2018_summarised <- summarise(realis2018_grouped, 
                          `Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE),
                          `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
                          `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE), 
                          `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```
## 16.3.4 Grouped summaries with the pipe

```{r}
realis2018_summarised <- realis2018 %>% 
  group_by(`Project Name`,`Planning Region`, 
           `Planning Area`, `Property Type`, 
           `Type of Sale`) %>%
  summarise(`Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE), 
            `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
            `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE),
            `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```
## 16.4.1 Designing a static treemap

```{r}
realis2018_selected <- realis2018_summarised %>%
  filter(`Property Type` == "Condominium", `Type of Sale` == "Resale")
```
## 16.4.2 Using the basic arguments

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```
## 16.4.3 Working with vColor and type arguments

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type = "value",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```
## 16.4.5 The “value” type treemap

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```
## 16.4.6 The “manual” type treemap

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```
## Ex single colour palette(Blue)

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```
## 16.4.8 Working with algorithm argument

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "squarified",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```
## 16.4.9 Using sortID

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "pivotSize",
        sortID = "Median Transacted Price",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```
## 16.5.1 Designing a basic treemap

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`),
       layout = "scol",
       start = "bottomleft") + 
  geom_treemap() +
  scale_fill_gradient(low = "light blue", high = "blue")
```
## 16.5.2 Defining hierarchy

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`),
       start = "topleft") + 
  geom_treemap()
```
## Group by Planning Area

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap()
```
## Adding boundary line

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap() +
  geom_treemap_subgroup2_border(colour = "gray40",
                                size = 2) +
  geom_treemap_subgroup_border(colour = "gray20")
```
## 16.6.1 Installing d3treeR package

```{r，eval=FALSE}
install.packages("devtools")
```

```{r,eval=FALSE}
library(devtools)
install_github("timelyportfolio/d3treeR")
```

```{r}
library(d3treeR)
```

## 16.6.2 Designing An Interactive Treemap
```{r}
tm <- treemap(realis2018_summarised,
        index=c("Planning Region", "Planning Area"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        title="Private Residential Property Sold, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

# Self Trying

## 16.3.3 Self Trying: Grouped summaries without the pipe

```{r}
realis2018_grouped <- group_by(
  realis2018,
  `Planning Region`,
  `Property Type`
)

realis2018_summarised <- summarise(
  realis2018_grouped,
  `Number of Transactions` = n(),
  `Total Units Sold` = sum(`No. of Units`, na.rm = TRUE),
  `Total Area (sqm)` = sum(`Area (sqm)`, na.rm = TRUE),
  `Average Unit Price ($ psm)` = mean(`Unit Price ($ psm)`, na.rm = TRUE),
  `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE),
  `Average Transacted Price ($)` = mean(`Transacted Price ($)`, na.rm = TRUE),
  `Median Transacted Price ($)` = median(`Transacted Price ($)`, na.rm = TRUE)
)
```

## 16.3.4 Self Trying: Select resale apartment records

```{r}
realis2018_summarised <- realis2018 %>%
  group_by(`Planning Region`, `Planning Area`, `Property Type`, `Type of Sale`) %>%
  summarise(
    `Number of Transactions` = n(),
    `Total Units Sold` = sum(`No. of Units`, na.rm = TRUE),
    `Total Area (sqm)` = sum(`Area (sqm)`, na.rm = TRUE),
    `Average Unit Price ($ psm)` = mean(`Unit Price ($ psm)`, na.rm = TRUE),
    `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE),
    `Average Transacted Price ($)` = mean(`Transacted Price ($)`, na.rm = TRUE),
    `Median Transacted Price ($)` = median(`Transacted Price ($)`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(`Total Units Sold`))
```

## 16.4.1 Self Trying: Select resale apartment records

```{r}
realis2018_selected <- realis2018_summarised %>%
  filter(`Property Type` == "Apartment",
         `Type of Sale` == "Resale")
```

## 16.4.2 Self Trying: Using the basic arguments

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area"),
  vSize = "Total Units Sold",
  vColor = "Median Unit Price ($ psm)",
  title = "Resale Apartment by Planning Region and Area, 2018",
  title.legend = "Median Unit Price (S$ per sq. m)"
)
```

## 16.4.3 Self Trying: Working with vColor and type arguments

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area"),
  vSize = "Total Units Sold",
  vColor = "Median Unit Price ($ psm)",
  type = "value",
  title = "Resale Apartment by Planning Region and Area, 2018",
  title.legend = "Median Unit Price (S$ per sq. m)"
)
```

## 16.4.5 Self Trying: The "value" type treemap

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area"),
  vSize = "Total Units Sold",
  vColor = "Median Unit Price ($ psm)",
  type = "value",
  palette = "YlGnBu",
  title = "Resale Apartment by Planning Region and Area, 2018",
  title.legend = "Median Unit Price (S$ per sq. m)"
)
```

## 16.4.6 Self Trying: The "manual" type treemap

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area"),
  vSize = "Total Units Sold",
  vColor = "Average Transacted Price ($)",
  type = "manual",
  palette = "YlGnBu",
  title = "Resale Apartment by Planning Region and Area, 2018",
  title.legend = "Average Transacted Price (S$)"
)
```

## Self Trying: The "manual" type treemap with Blues palette

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area"),
  vSize = "Total Units Sold",
  vColor = "Average Transacted Price ($)",
  type = "manual",
  palette = "Blues",
  title = "Resale Apartment by Planning Region and Area, 2018",
  title.legend = "Average Transacted Price (S$)"
)
```


## 16.4.8 Self Trying: Working with algorithm argument

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area"),
  vSize = "Total Units Sold",
  vColor = "Average Transacted Price ($)",
  type = "manual",
  palette = "Blues",
  algorithm = "squarified",
  title = "Resale Apartment by Planning Region and Area, 2018",
  title.legend = "Average Transacted Price (S$)"
)
```

## 16.4.9 Self Trying: Using sortID

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area"),
  vSize = "Total Units Sold",
  vColor = "Average Transacted Price ($)",
  type = "manual",
  palette = "Blues",
  algorithm = "pivotSize",
  sortID = "Median Transacted Price ($)",
  title = "Resale Apartment by Planning Region and Area, 2018",
  title.legend = "Average Transacted Price (S$)"
)
```

## 16.5.1 Self Trying: Designing a basic treemap

```{r}
ggplot(
  data = realis2018_selected,
  aes(
    area = `Total Units Sold`,
    fill = `Average Transacted Price ($)`
  )
) +
  geom_treemap(
    layout = "scol",
    start = "bottomleft"
  ) +
  scale_fill_gradient(
    low = "light blue",
    high = "blue"
  ) +
  labs(
    title = "Basic Treemap of Resale Apartment Transactions in 2018",
    fill = "Average Transacted Price"
  )
```

## 16.5.2 Self Trying: Defining hierarchy

```{r}
ggplot(
  data = realis2018_selected,
  aes(
    area = `Total Units Sold`,
    fill = `Average Transacted Price ($)`,
    subgroup = `Planning Region`
  )
) +
  geom_treemap(
    layout = "scol",
    start = "topleft"
  ) +
  geom_treemap_subgroup_border(
    colour = "white",
    size = 2
  ) +
  scale_fill_gradient(
    low = "light blue",
    high = "blue"
  ) +
  labs(
    title = "Treemap of Resale Apartment Transactions by Planning Region",
    fill = "Average Transacted Price"
  )
```

## Self Trying: Group by Planning Area

```{r}
ggplot(
  data = realis2018_selected,
  aes(
    area = `Total Units Sold`,
    fill = `Average Transacted Price ($)`,
    subgroup = `Planning Region`,
    subgroup2 = `Planning Area`
  )
) +
  geom_treemap(
    layout = "scol",
    start = "topleft"
  ) +
  geom_treemap_subgroup_border(
    colour = "white",
    size = 2
  ) +
  geom_treemap_subgroup2_border(
    colour = "grey70",
    size = 0.8
  ) +
  scale_fill_gradient(
    low = "light blue",
    high = "blue"
  ) +
  labs(
    title = "Treemap of Resale Apartment Transactions by Planning Region and Area",
    fill = "Average Transacted Price"
  )
```

### Self Trying： Adding boundary line

```{r}
ggplot(
  data = realis2018_selected,
  aes(
    area = `Total Units Sold`,
    fill = `Average Transacted Price ($)`,
    subgroup = `Planning Region`,
    subgroup2 = `Planning Area`
  )
) +
  geom_treemap(
    layout = "scol",
    start = "topleft"
  ) +
  geom_treemap_subgroup2_border(
    colour = "gray70",
    size = 1
  ) +
  geom_treemap_subgroup_border(
    colour = "gray20",
    size = 2
  ) +
  scale_fill_gradient(
    low = "light blue",
    high = "blue"
  ) +
  labs(
    title = "Treemap of Resale Apartment Transactions by Planning Region and Area",
    fill = "Average Transacted Price"
  )
```
### Self Trying: Designing An Interactive Treemap
```{r}
tm <- treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area"),
  vSize = "Total Units Sold",
  vColor = "Average Transacted Price ($)",
  type = "value",
  palette = "Blues",
  title = "Interactive Treemap of Resale Apartment Transactions, 2018",
  title.legend = "Average Transacted Price (S$)"
)
```

## Then d3tree() is used to build an interactive treemap
```{r}
d3tree(tm,rootname = "Singapore" )
```


