---
title: "Hands-on Exercise 9a: Choropleth Mapping with R"
author: "JUNCHEN LI"
date: "June 17, 2026"
date-modified: "last-modified"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
execute:
  warning: false
  message: false
---

# Overview

This hands-on exercise follows the learning flow from Chapter 21 of *R for Visual Analytics*. The teacher's example focuses on choropleth mapping with **tmap** by combining Singapore planning subzone boundary data with population attribute data. After reproducing the main classroom examples, I add a **Self Trying** section to explore ageing and dependency patterns in a slightly different way.

# 1 Getting Started

## 1.1 Load R packages

```{r}
pacman::p_load(sf, tmap, tidyverse, readr, DT)
```

For this exercise, `sf` is used to handle geospatial data, `tmap` is used to draw thematic maps, and `tidyverse` is used for data wrangling.

# 2 Importing Data into R

## 2.1 Importing geospatial data

The geospatial data is the Master Plan 2014 Subzone Boundary shapefile. It should be stored in the folder `data/geospatial`.

```{r}
mpsz <- st_read(
  dsn = "data/geospatial",
  layer = "MP14_SUBZONE_WEB_PL"
)
```

```{r}
mpsz
```

## 2.2 Importing attribute data

The attribute data is the Singapore Residents by Planning Area, Subzone, Age Group, Sex and Type of Dwelling dataset from 2011 to 2020. It should be stored in the folder `data/aspatial`.

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

```{r}
glimpse(popdata)
```

# 3 Data Preparation

The teacher's example prepares a 2020 population table at the planning subzone level. The original learning point is to derive the number of young residents, economically active residents, aged residents, total residents, and the dependency ratio.

## 3.1 Creating a reusable population preparation function

Instead of relying on column positions after `pivot_wider()`, I use a more robust method. The starting age is extracted from the age-group label and then grouped into `YOUNG`, `ECONOMY ACTIVE`, and `AGED`.

```{r}
prepare_population <- function(selected_year) {
  popdata %>%
    filter(Time == selected_year) %>%
    mutate(
      age_start = parse_number(AG),
      age_group_type = case_when(
        age_start <= 24 ~ "YOUNG",
        age_start >= 25 & age_start <= 64 ~ "ECONOMY ACTIVE",
        age_start >= 65 ~ "AGED",
        TRUE ~ NA_character_
      )
    ) %>%
    filter(!is.na(age_group_type)) %>%
    group_by(PA, SZ, age_group_type) %>%
    summarise(POP = sum(Pop, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(
      names_from = age_group_type,
      values_from = POP,
      values_fill = 0
    ) %>%
    mutate(
      TOTAL = YOUNG + `ECONOMY ACTIVE` + AGED,
      DEPENDENCY = (YOUNG + AGED) / `ECONOMY ACTIVE`,
      YOUNG_SHARE = YOUNG / TOTAL,
      AGED_SHARE = AGED / TOTAL,
      WORKING_SHARE = `ECONOMY ACTIVE` / TOTAL
    ) %>%
    select(
      PA, SZ, YOUNG, `ECONOMY ACTIVE`, AGED, TOTAL,
      DEPENDENCY, YOUNG_SHARE, AGED_SHARE, WORKING_SHARE
    ) %>%
    mutate(across(c(PA, SZ), toupper)) %>%
    filter(`ECONOMY ACTIVE` > 0)
}
```

## 3.2 Preparing the 2020 table

```{r}
popdata2020 <- prepare_population(2020)
```

```{r}
glimpse(popdata2020)
```

## 3.3 Joining attribute data with geospatial data

The georelational join uses `SUBZONE_N` from the geospatial layer and `SZ` from the attribute table.

```{r}
mpsz_pop2020 <- left_join(
  mpsz,
  popdata2020,
  by = c("SUBZONE_N" = "SZ")
)
```

```{r}
dir.create("data/rds", recursive = TRUE, showWarnings = FALSE)
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

# 4 Teacher Example: Choropleth Mapping with tmap

## 4.1 Plotting a choropleth map quickly with qtm()

The simplest way to create a choropleth map is to use `qtm()`.

::: panel-tabset
### The plot

```{r, echo=FALSE}
tmap_mode("plot")
qtm(
  mpsz_pop2020,
  fill = "DEPENDENCY"
)
```

### The code

```{r, eval=FALSE}
tmap_mode("plot")
qtm(
  mpsz_pop2020,
  fill = "DEPENDENCY"
)
```
:::

## 4.2 Creating a choropleth map with tm_shape() and tm_polygons()

The `tm_shape()` and `tm_polygons()` approach provides more control over classification, colour, legend, title and borders.

::: panel-tabset
### The plot

```{r, echo=FALSE}
tm_shape(mpsz_pop2020) +
  tm_polygons(
    "DEPENDENCY",
    style = "quantile",
    n = 5,
    palette = "Blues",
    title = "Dependency ratio"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(
    main.title = "Distribution of Dependency Ratio by Planning Subzone",
    frame = TRUE
  )
```

### The code

```{r, eval=FALSE}
tm_shape(mpsz_pop2020) +
  tm_polygons(
    "DEPENDENCY",
    style = "quantile",
    n = 5,
    palette = "Blues",
    title = "Dependency ratio"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(
    main.title = "Distribution of Dependency Ratio by Planning Subzone",
    frame = TRUE
  )
```
:::

## 4.3 Drawing a base map

A base map helps us check whether the spatial layer is imported correctly.

::: panel-tabset
### The plot

```{r, echo=FALSE}
tm_shape(mpsz_pop2020) +
  tm_polygons(col = "white", border.col = "grey60") +
  tm_layout(main.title = "Planning Subzone Boundary")
```

### The code

```{r, eval=FALSE}
tm_shape(mpsz_pop2020) +
  tm_polygons(col = "white", border.col = "grey60") +
  tm_layout(main.title = "Planning Subzone Boundary")
```
:::

## 4.4 Drawing choropleth map with tm_fill() and tm_borders()

`tm_fill()` shades the polygon, while `tm_borders()` draws the boundary line.

::: panel-tabset
### The plot

```{r, echo=FALSE}
tm_shape(mpsz_pop2020) +
  tm_fill(
    "DEPENDENCY",
    style = "pretty",
    palette = "YlOrRd",
    title = "Dependency ratio"
  ) +
  tm_borders(lwd = 0.01, alpha = 0.5) +
  tm_layout(main.title = "Dependency Ratio with Subzone Borders")
```

### The code

```{r, eval=FALSE}
tm_shape(mpsz_pop2020) +
  tm_fill(
    "DEPENDENCY",
    style = "pretty",
    palette = "YlOrRd",
    title = "Dependency ratio"
  ) +
  tm_borders(lwd = 0.01, alpha = 0.5) +
  tm_layout(main.title = "Dependency Ratio with Subzone Borders")
```
:::

## 4.5 Comparing data classification methods

Different classification methods can lead to different visual interpretations. Here, Jenks and equal interval are compared.

### 4.5.1 Jenks classification

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons(
    "DEPENDENCY",
    style = "jenks",
    n = 5,
    palette = "Blues",
    title = "Dependency ratio"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Jenks Classification")
```

### 4.5.2 Equal interval classification

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons(
    "DEPENDENCY",
    style = "equal",
    n = 5,
    palette = "Blues",
    title = "Dependency ratio"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal Interval Classification")
```

## 4.6 Choropleth map with custom breaks

Before using custom breaks, it is useful to check descriptive statistics.

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons(
    "DEPENDENCY",
    breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00, Inf),
    palette = "YlOrRd",
    title = "Dependency ratio"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Dependency Ratio with Custom Breaks")
```

## 4.7 Colour scheme

### 4.7.1 Using a green colour palette

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons(
    "DEPENDENCY",
    style = "quantile",
    n = 5,
    palette = "Greens",
    title = "Dependency ratio"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Dependency Ratio with Green Palette")
```

### 4.7.2 Reversing the colour palette

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons(
    "DEPENDENCY",
    style = "quantile",
    n = 5,
    palette = "-Greens",
    title = "Dependency ratio"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Dependency Ratio with Reversed Green Palette")
```

## 4.8 Map layout and cartographic elements

This version adds title, compass, scale bar and credits.

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons(
    fill = "DEPENDENCY",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "Blues"
    ),
    fill.legend = tm_legend(
      title = "Dependency ratio"
    )
  ) +
  tm_borders(
    col = "grey40",
    lwd = 0.5
  ) +
  tm_compass(
    type = "8star",
    size = 2,
    position = c("right", "top")
  ) +
  tm_scalebar(
    position = c("left", "bottom")
  ) +
  tm_layout(
    main.title = "Distribution of Dependency Ratio by Planning Subzone",
    frame = TRUE
  ) +
  tm_credits(
    "Source: URA Master Plan 2014 Subzone Boundary and DOS population data",
    position = c("left", "bottom")
  )
```

## 4.9 Small multiple choropleth maps

Small multiples allow several variables to be compared side by side.

### 4.9.1 Mapping young and aged population counts

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill(
    c("YOUNG", "AGED"),
    style = "quantile",
    palette = "Blues"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "bottom"))
```

### 4.9.2 Mapping by region with facets

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill(
    "DEPENDENCY",
    style = "quantile",
    palette = "Blues"
  ) +
  tm_facets(by = "REGION_N", free.coords = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_layout(
    legend.show = FALSE,
    main.title = "Dependency Ratio by Region"
  )
```

### 4.9.3 Arranging multiple standalone maps

```{r}
youngmap <- tm_shape(mpsz_pop2020) +
  tm_polygons(
    "YOUNG",
    style = "quantile",
    palette = "Blues",
    title = "Young population"
  ) +
  tm_borders(alpha = 0.5)

agedmap <- tm_shape(mpsz_pop2020) +
  tm_polygons(
    "AGED",
    style = "quantile",
    palette = "Greens",
    title = "Aged population"
  ) +
  tm_borders(alpha = 0.5)

tmap_arrange(youngmap, agedmap, ncol = 2)
```

## 4.10 Mapping selected spatial objects

This map focuses only on subzones located in the Central Region.

```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N == "CENTRAL REGION", ]) +
  tm_fill(
    "DEPENDENCY",
    style = "quantile",
    palette = "Blues",
    title = "Dependency ratio"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(
    legend.outside = TRUE,
    frame = FALSE,
    main.title = "Dependency Ratio in the Central Region"
  )
```

# 5 Self Trying: Exploring Ageing and Dependency Patterns

In this self-trying section, I use the same choropleth mapping techniques from the teacher's example, but the analytical focus is slightly different. Instead of only mapping the dependency ratio, I compare ageing share, young share, working-age share and the change in dependency ratio between 2011 and 2020.

## 5.1 Preparing additional indicators for 2020

```{r}
self_2020 <- mpsz_pop2020 %>%
  mutate(
    DEPENDENCY_LABEL = case_when(
      DEPENDENCY < 0.60 ~ "Low dependency",
      DEPENDENCY >= 0.60 & DEPENDENCY < 0.80 ~ "Moderate dependency",
      DEPENDENCY >= 0.80 & DEPENDENCY < 1.00 ~ "High dependency",
      DEPENDENCY >= 1.00 ~ "Very high dependency",
      TRUE ~ NA_character_
    )
  )
```

```{r}
self_2020 %>%
  st_drop_geometry() %>%
  summarise(
    min_dependency = min(DEPENDENCY, na.rm = TRUE),
    median_dependency = median(DEPENDENCY, na.rm = TRUE),
    mean_dependency = mean(DEPENDENCY, na.rm = TRUE),
    max_dependency = max(DEPENDENCY, na.rm = TRUE),
    min_aged_share = min(AGED_SHARE, na.rm = TRUE),
    max_aged_share = max(AGED_SHARE, na.rm = TRUE)
  )
```

## 5.2 Mapping aged share by planning subzone

This map focuses on where aged residents form a larger share of the local population.

::: panel-tabset
### The plot

```{r, echo=FALSE}
tm_shape(self_2020) +
  tm_polygons(
    "AGED_SHARE",
    style = "jenks",
    n = 5,
    palette = "Reds",
    title = "Aged share"
  ) +
  tm_borders(alpha = 0.4) +
  tm_layout(
    main.title = "Share of Aged Residents by Planning Subzone, 2020",
    legend.outside = TRUE
  )
```

### The code

```{r, eval=FALSE}
tm_shape(self_2020) +
  tm_polygons(
    "AGED_SHARE",
    style = "jenks",
    n = 5,
    palette = "Reds",
    title = "Aged share"
  ) +
  tm_borders(alpha = 0.4) +
  tm_layout(
    main.title = "Share of Aged Residents by Planning Subzone, 2020",
    legend.outside = TRUE
  )
```
:::

## 5.3 Comparing dependency, young share and aged share

This small multiple map compares three related population indicators. It helps reveal whether high dependency is mainly driven by younger residents or older residents.

```{r}
tm_shape(self_2020) +
  tm_polygons(
    c("DEPENDENCY", "YOUNG_SHARE", "AGED_SHARE"),
    style = c("quantile", "quantile", "quantile"),
    palette = list("Purples", "Blues", "Reds")
  ) +
  tm_borders(alpha = 0.4) +
  tm_layout(
    legend.position = c("right", "bottom"),
    main.title = "Comparing Dependency, Young Share and Aged Share"
  )
```

## 5.4 Mapping high-dependency subzones

Here, I highlight subzones where dependency ratio is at least 1. This means the number of young and aged residents is at least as large as the number of economically active residents.

```{r}
high_dependency <- self_2020 %>%
  filter(DEPENDENCY >= 1)
```

```{r}
tm_shape(self_2020) +
  tm_polygons(col = "grey90", border.col = "white") +
  tm_shape(high_dependency) +
  tm_polygons(
    "DEPENDENCY",
    style = "quantile",
    n = 5,
    palette = "Reds",
    title = "Dependency ratio"
  ) +
  tm_borders(col = "grey30", lwd = 0.2) +
  tm_layout(
    main.title = "Subzones with Very High Dependency Ratio, 2020",
    legend.outside = TRUE
  )
```

## 5.5 Listing the top high-dependency subzones

A table is useful because a map shows location well, but it does not always allow precise ranking.

```{r}
self_2020 %>%
  st_drop_geometry() %>%
  select(PLN_AREA_N, SUBZONE_N, REGION_N, DEPENDENCY, YOUNG_SHARE, AGED_SHARE, TOTAL) %>%
  arrange(desc(DEPENDENCY)) %>%
  slice_head(n = 20) %>%
  DT::datatable(
    filter = "top",
    rownames = FALSE,
    options = list(pageLength = 10, autoWidth = TRUE),
    colnames = c(
      "Planning Area", "Subzone", "Region", "Dependency Ratio",
      "Young Share", "Aged Share", "Total Population"
    )
  )
```

## 5.6 Preparing 2011 and 2020 dependency change

This part extends the teacher's example by comparing two years. The aim is to identify subzones where the dependency ratio increased or decreased between 2011 and 2020.

```{r}
popdata2011 <- prepare_population(2011) %>%
  select(SZ, DEPENDENCY_2011 = DEPENDENCY)

popdata2020_change <- prepare_population(2020) %>%
  select(SZ, DEPENDENCY_2020 = DEPENDENCY)

dependency_change <- popdata2011 %>%
  inner_join(popdata2020_change, by = "SZ") %>%
  mutate(
    DEPENDENCY_CHANGE = DEPENDENCY_2020 - DEPENDENCY_2011
  )

mpsz_dependency_change <- left_join(
  mpsz,
  dependency_change,
  by = c("SUBZONE_N" = "SZ")
)
```

## 5.7 Mapping dependency change from 2011 to 2020

A diverging colour palette is used because the value can be positive or negative.

```{r}
tm_shape(mpsz_dependency_change) +
  tm_polygons(
    "DEPENDENCY_CHANGE",
    style = "pretty",
    palette = "RdBu",
    title = "Change in dependency ratio"
  ) +
  tm_borders(alpha = 0.4) +
  tm_layout(
    main.title = "Change in Dependency Ratio by Planning Subzone, 2011-2020",
    legend.outside = TRUE
  )
```

## 5.8 Small multiple maps for 2011 and 2020

This final map compares the spatial distribution of dependency ratio in 2011 and 2020 side by side.

```{r}
mpsz_pop2011 <- left_join(
  mpsz,
  prepare_population(2011),
  by = c("SUBZONE_N" = "SZ")
)

map_2011 <- tm_shape(mpsz_pop2011) +
  tm_polygons(
    "DEPENDENCY",
    style = "quantile",
    n = 5,
    palette = "Blues",
    title = "2011"
  ) +
  tm_borders(alpha = 0.4) +
  tm_layout(main.title = "2011")

map_2020 <- tm_shape(mpsz_pop2020) +
  tm_polygons(
    "DEPENDENCY",
    style = "quantile",
    n = 5,
    palette = "Blues",
    title = "2020"
  ) +
  tm_borders(alpha = 0.4) +
  tm_layout(main.title = "2020")

tmap_arrange(map_2011, map_2020, ncol = 2)
```

## 5.9 Interpretation

The self-trying maps show that dependency ratio is not only a single demographic measure; it can be decomposed into young share and aged share. Some subzones may show high dependency because they have many young residents, while others may show high dependency because they have a larger aged population. The 2011-2020 change map provides an additional time dimension and helps identify subzones where demographic pressure has increased over time. This makes the choropleth analysis more useful for planning because it shows both spatial pattern and temporal change.
