---
title: "Hands-on Exercise 09c: Analytical Mapping"
author: "JUNCHEN LI"
date: "June 15, 2026"
date-modified: "last-modified"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
execute:
  warning: false
  message: false
  freeze: false
---

# Overview

This hands-on exercise follows Chapter 23, **Analytical Mapping**. The teacher example focuses on analytical maps for Nigeria water points at the LGA level. The main techniques include choropleth mapping, rate mapping, percentile maps, and box maps.

The self-trying section keeps the same learning flow, but focuses on a slightly different analytical question:

> Which LGAs should be considered higher priority when both the number and the proportion of non-functional water points are considered?

# 1 Getting Started

## 1.1 Installing and loading packages

```{r}
pacman::p_load(tmap, tidyverse, sf, DT)
```

```{r}
tmap_mode("plot")
```

# 2 Teacher Example: Importing Data

## 2.1 Importing prepared geospatial data

The prepared data set `NGA_wp.rds` is a simple feature polygon data frame. It contains water point information at the Local Government Area (LGA) level in Nigeria.

```{r}
NGA_wp <- read_rds("data/rds/NGA_wp.rds")
```

```{r}
NGA_wp
```

```{r}
glimpse(NGA_wp)
```

# 3 Teacher Example: Basic Choropleth Mapping

## 3.1 Visualising the distribution of functional water points

This map shows the number of functional water points by LGA. It is useful as a basic count map, but it should be interpreted carefully because larger or more populated LGAs may naturally have more water points.

```{r}
p_functional <- tm_shape(NGA_wp) +
  tm_polygons(
    fill = "wp_functional",
    fill.scale = tm_scale_intervals(
      style = "equal",
      n = 10,
      values = "brewer.blues"
    ),
    fill.legend = tm_legend(
      title = "Functional water points",
      position = c("right", "bottom")
    )
  ) +
  tm_borders(
    lwd = 0.1,
    fill_alpha = 1
  ) +
  tm_title("Distribution of Functional Water Points by LGAs")

p_functional
```

## 3.2 Visualising the distribution of total water points

```{r}
p_total <- tm_shape(NGA_wp) +
  tm_polygons(
    fill = "total_wp",
    fill.scale = tm_scale_intervals(
      style = "equal",
      n = 10,
      values = "brewer.blues"
    ),
    fill.legend = tm_legend(
      title = "Total water points",
      position = c("right", "bottom")
    )
  ) +
  tm_borders(
    lwd = 0.1,
    fill_alpha = 1
  ) +
  tm_title("Distribution of Total Water Points by LGAs")

p_total
```

## 3.3 Comparing total and functional water points

```{r}
tmap_arrange(p_total, p_functional, nrow = 1)
```

# 4 Teacher Example: Choropleth Map for Rates

## 4.1 Deriving rate variables

A count map can be misleading because water points are not equally distributed across LGAs. Therefore, it is useful to derive rates.

```{r}
NGA_wp <- NGA_wp %>%
  mutate(
    pct_functional = if_else(
      total_wp > 0,
      wp_functional / total_wp,
      NA_real_
    ),
    pct_nonfunctional = if_else(
      total_wp > 0,
      wp_nonfunctional / total_wp,
      NA_real_
    )
  )
```

```{r}
summary(NGA_wp$pct_functional)
summary(NGA_wp$pct_nonfunctional)
```

## 4.2 Plotting a rate map of functional water points

```{r}
tm_shape(NGA_wp) +
  tm_polygons(
    fill = "pct_functional",
    fill.scale = tm_scale_intervals(
      style = "equal",
      n = 10,
      values = "brewer.blues"
    ),
    fill.legend = tm_legend(
      title = "Functional rate",
      position = c("right", "bottom")
    )
  ) +
  tm_borders(
    lwd = 0.1,
    fill_alpha = 1
  ) +
  tm_title("Rate Map of Functional Water Points by LGAs")
```

## 4.3 Plotting a rate map of non-functional water points

```{r}
tm_shape(NGA_wp) +
  tm_polygons(
    fill = "pct_nonfunctional",
    fill.scale = tm_scale_intervals(
      style = "equal",
      n = 10,
      values = "brewer.reds"
    ),
    fill.legend = tm_legend(
      title = "Non-functional rate",
      position = c("right", "bottom")
    )
  ) +
  tm_borders(
    lwd = 0.1,
    fill_alpha = 1
  ) +
  tm_title("Rate Map of Non-functional Water Points by LGAs")
```

# 5 Teacher Example: Extreme Value Maps

Extreme value maps help identify spatial outliers. Instead of focusing only on general distribution, they highlight LGAs located at the lower or upper tails of the distribution.

## 5.1 Preparing clean data

```{r}
NGA_wp_clean <- NGA_wp %>%
  drop_na(pct_functional, pct_nonfunctional, wp_nonfunctional, total_wp)
```

## 5.2 Creating a helper function to extract a variable

```{r}
get_var <- function(vname, df) {
  v <- df[vname] %>%
    st_set_geometry(NULL)
  v <- unname(v[, 1])
  return(v)
}
```

## 5.3 Percentile map function

The percentile map uses six categories: 0-1%, 1-10%, 10-50%, 50-90%, 90-99%, and 99-100%.

```{r}
percentmap <- function(vnam, df, legtitle = NA, mtitle = "Percentile Map",
                       palette = "brewer.reds") {
  
  percent <- c(0, .01, .1, .5, .9, .99, 1)
  
  var <- get_var(vnam, df)
  var <- var[!is.na(var)]
  
  bperc <- quantile(
    var,
    percent,
    na.rm = TRUE,
    names = FALSE
  )
  
  bperc <- unique(as.numeric(bperc))
  
  if (length(bperc) < 4) {
    bperc <- pretty(var, n = 5)
    bperc <- unique(as.numeric(bperc))
  }
  
  tm_shape(df) +
    tm_polygons(
      fill = vnam,
      fill.scale = tm_scale_intervals(
        breaks = bperc,
        values = palette
      ),
      fill.legend = tm_legend(
        title = legtitle,
        position = c("right", "bottom")
      )
    ) +
    tm_borders(lwd = 0.1) +
    tm_title(mtitle)
}
```

## 5.4 Test driving the percentile map

```{r}
percentmap(
  "total_wp",
  NGA_wp_clean,
  legtitle = "Total water points",
  mtitle = "Percentile Map of Total Water Points"
)
```

## 5.5 Using a boxplot to understand outliers

```{r}
ggplot(
  data = NGA_wp_clean,
  aes(
    x = "",
    y = wp_nonfunctional
  )
) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Non-functional Water Points",
    x = NULL,
    y = "Number of non-functional water points"
  ) +
  theme_minimal()
```

## 5.6 Creating the boxbreaks function

```{r}
boxbreaks <- function(v, mult = 1.5) {

  qv <- unname(quantile(v, na.rm = TRUE))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr

  bb <- vector(mode = "numeric", length = 7)

  if (lofence < qv[1]) {
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }

  if (upfence > qv[5]) {
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }

  bb[3:5] <- qv[2:4]

  return(bb)
}
```

```{r}
var_nonfunctional <- get_var("wp_nonfunctional", NGA_wp_clean)
boxbreaks(var_nonfunctional)
```

## 5.7 Box map function

```{r}
boxmap <- function(vnam, df,
                   legtitle = NA,
                   mtitle = "Box Map",
                   mult = 1.5) {

  var <- get_var(vnam, df)
  bb <- boxbreaks(var, mult)

  tm_shape(df) +
    tm_polygons(
      fill = vnam,
      fill.scale = tm_scale_intervals(
        breaks = bb,
        values = "brewer.blues"
      ),
      fill.legend = tm_legend(
        title = legtitle,
        position = c("right", "bottom")
      )
    ) +
    tm_borders(lwd = 0.1) +
    tm_title(mtitle)
}
```

## 5.8 Test driving the box map

```{r}
boxmap(
  "wp_nonfunctional",
  NGA_wp_clean,
  legtitle = "Non-functional water points",
  mtitle = "Box Map of Non-functional Water Points"
)
```

# 6 Self Trying: Mapping Water Point Service Risk

In this self-trying section, I use the same analytical mapping techniques from the teacher example, but I focus on a practical planning question. Instead of mapping only the number of water points or only the rate of functionality, I combine both the number and the rate of non-functional water points to identify higher-priority LGAs.

## 6.1 Preparing additional indicators

```{r}
NGA_self <- NGA_wp %>%
  mutate(
    pct_functional = if_else(
      total_wp > 0,
      wp_functional / total_wp,
      NA_real_
    ),
    pct_nonfunctional = if_else(
      total_wp > 0,
      wp_nonfunctional / total_wp,
      NA_real_
    )
  ) %>%
  drop_na(pct_functional, pct_nonfunctional, wp_nonfunctional, total_wp)
```

The following code tries to identify a suitable LGA name column. If the data set uses a different name field, the code still creates a usable `lga_label`.

```{r}
candidate_name_cols <- c(
  "shapeName", "ADM2_EN", "LGA", "lga", "lga_name",
  "ADM2_NAME", "name", "Name", "NAME"
)

lga_name_col <- candidate_name_cols[
  candidate_name_cols %in% names(NGA_self)
][1]

if (is.na(lga_name_col)) {
  NGA_self <- NGA_self %>%
    mutate(lga_label = paste0("LGA_", row_number()))
} else {
  NGA_self <- NGA_self %>%
    mutate(lga_label = as.character(.data[[lga_name_col]]))
}
```

## 6.2 Comparing count map and rate map

This comparison shows why rate maps are important. A high count of non-functional water points may simply reflect a large number of total water points, while a high rate indicates a larger share of local water points are non-functional.

```{r}
p_count_nonfunctional <- tm_shape(NGA_self) +
  tm_polygons(
    fill = "wp_nonfunctional",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.reds"
    ),
    fill.legend = tm_legend(
      title = "Non-functional count",
      position = c("right", "bottom")
    )
  ) +
  tm_borders(lwd = 0.1) +
  tm_title("Count of Non-functional Water Points")

p_rate_nonfunctional <- tm_shape(NGA_self) +
  tm_polygons(
    fill = "pct_nonfunctional",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 5,
      values = "brewer.reds"
    ),
    fill.legend = tm_legend(
      title = "Non-functional rate",
      position = c("right", "bottom")
    )
  ) +
  tm_borders(lwd = 0.1) +
  tm_title("Rate of Non-functional Water Points")

tmap_arrange(
  p_count_nonfunctional,
  p_rate_nonfunctional,
  nrow = 1
)
```

## 6.3 Creating a priority flag

I define priority LGAs as areas that are in the top 10% for both the number of non-functional water points and the proportion of non-functional water points. This rule helps avoid overreacting to LGAs with a high rate but very few water points, or LGAs with many failures simply because they have many water points.

```{r}
count_cutoff <- quantile(
  NGA_self$wp_nonfunctional,
  0.90,
  na.rm = TRUE
)

rate_cutoff <- quantile(
  NGA_self$pct_nonfunctional,
  0.90,
  na.rm = TRUE
)

NGA_self <- NGA_self %>%
  mutate(
    high_count = wp_nonfunctional >= count_cutoff,
    high_rate = pct_nonfunctional >= rate_cutoff,
    priority_group = case_when(
      high_count & high_rate ~ "High count and high rate",
      high_count & !high_rate ~ "High count only",
      !high_count & high_rate ~ "High rate only",
      TRUE ~ "Lower priority"
    ),
    priority_group = factor(
      priority_group,
      levels = c(
        "High count and high rate",
        "High count only",
        "High rate only",
        "Lower priority"
      )
    )
  )
```

```{r}
NGA_self %>%
  st_set_geometry(NULL) %>%
  count(priority_group)
```

## 6.4 Mapping priority groups

```{r}
tm_shape(NGA_self) +
  tm_polygons(
    fill = "priority_group",
    fill.scale = tm_scale_categorical(
      values = c(
        "High count and high rate" = "#b2182b",
        "High count only" = "#ef8a62",
        "High rate only" = "#fddbc7",
        "Lower priority" = "#f0f0f0"
      )
    ),
    fill.legend = tm_legend(
      title = "Priority group",
      position = c("right", "bottom")
    )
  ) +
  tm_borders(lwd = 0.1) +
  tm_title("Priority Areas Based on Non-functional Water Point Burden")
```

## 6.5 Percentile map of non-functional rate

```{r}
percentmap(
  "pct_nonfunctional",
  NGA_self,
  legtitle = "Non-functional rate",
  mtitle = "Percentile Map of Non-functional Water Point Rate",
  palette = "brewer.reds"
)
```

## 6.6 Box map of non-functional rate

```{r}
boxmap(
  "pct_nonfunctional",
  NGA_self,
  legtitle = "Non-functional rate",
  mtitle = "Box Map of Non-functional Water Point Rate"
)
```

## 6.7 Top priority LGAs table

```{r}
priority_table <- NGA_self %>%
  st_set_geometry(NULL) %>%
  transmute(
    LGA = lga_label,
    total_wp,
    wp_functional,
    wp_nonfunctional,
    pct_functional = round(pct_functional, 3),
    pct_nonfunctional = round(pct_nonfunctional, 3),
    priority_group
  ) %>%
  arrange(
    desc(priority_group == "High count and high rate"),
    desc(wp_nonfunctional),
    desc(pct_nonfunctional)
  )

DT::datatable(
  priority_table,
  filter = "top",
  rownames = FALSE,
  options = list(
    pageLength = 15,
    autoWidth = TRUE
  )
)
```

## 6.8 Focused map of the most critical LGAs

This final map keeps all LGAs visible but highlights the areas that are both high in count and high in rate. It supports a clearer planning story.

```{r}
critical_lgas <- NGA_self %>%
  filter(priority_group == "High count and high rate")

tm_shape(NGA_self) +
  tm_polygons(
    col = "grey80",
    fill = "grey95"
  ) +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(critical_lgas) +
  tm_polygons(
    fill = "pct_nonfunctional",
    fill.scale = tm_scale_intervals(
      style = "quantile",
      n = 4,
      values = "brewer.reds"
    ),
    fill.legend = tm_legend(
      title = "Non-functional rate",
      position = c("right", "bottom")
    )
  ) +
  tm_borders(lwd = 0.3, col = "black") +
  tm_title("Critical LGAs: High Count and High Rate of Non-functional Water Points")
```

## 6.9 Interpretation

The self-trying analysis shows that count maps and rate maps answer different questions. Count maps help identify LGAs with a large absolute number of non-functional water points, while rate maps reveal where non-functionality is relatively more severe. The priority map combines both perspectives and highlights LGAs that are likely to require more urgent intervention because they have both a high number and a high proportion of non-functional water points. This makes the map more useful for planning than using only one indicator.
