---
title: "Executive Summary"
subtitle: "Visual Analytics of Motor Insurance Portfolio"
author: "JUNCHEN LI"
date: "May 24, 2026"
date-modified: "last-modified"
format:
  revealjs:
    theme: dark
    slide-number: true
    transition: fade
    width: 1600
    height: 900
    code-fold: true
editor: visual
execute:
  eval: true
  echo: false
  warning: false
  message: false
  freeze: auto
---

```{r}
pacman::p_load(
  tidyverse,
  janitor,
  scales,
  ggthemes,
  plotly,
  broom
)

theme_set(theme_minimal())
```

```{r}
motor_raw <- read_delim(
  "data/Dataset of motor insurance portfolio.csv",
  delim = ";",
  show_col_types = FALSE
)

motor <- motor_raw %>%
  clean_names()

motor_clean <- motor %>%
  drop_na(
    fuel_type,
    vehicle_value,
    vehicle_age,
    age_driving_licence
  ) %>%
  mutate(
    claim_indicator = if_else(total_claims > 0, 1, 0),
    profit = total_premium - total_incurred,
    loss_ratio = if_else(total_premium > 0, total_incurred / total_premium, NA_real_),
    claim_frequency = if_else(total_exposure > 0, total_claims / total_exposure, NA_real_),
    claim_severity = if_else(total_claims > 0, total_incurred / total_claims, NA_real_),
    driver_age_group = cut(
      driver_age,
      breaks = c(17, 25, 35, 45, 55, 65, Inf),
      labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+"),
      right = TRUE
    ),
    vehicle_age_group = cut(
      vehicle_age,
      breaks = c(-1, 5, 10, 20, 30, Inf),
      labels = c("0-5", "6-10", "11-20", "21-30", "31+"),
      right = TRUE
    )
  )
```

```{r}
portfolio_summary <- motor_clean %>%
  summarise(
    number_of_records = n(),
    unique_insured = n_distinct(insured_id),
    total_premium = sum(total_premium, na.rm = TRUE),
    total_incurred = sum(total_incurred, na.rm = TRUE),
    total_claims = sum(total_claims, na.rm = TRUE),
    total_exposure = sum(total_exposure, na.rm = TRUE),
    overall_profit = total_premium - total_incurred,
    overall_loss_ratio = total_incurred / total_premium,
    overall_claim_frequency = total_claims / total_exposure,
    overall_claim_severity = total_incurred / total_claims
  )

year_summary <- motor_clean %>%
  group_by(year) %>%
  summarise(
    premium = sum(total_premium, na.rm = TRUE),
    incurred = sum(total_incurred, na.rm = TRUE),
    claims = sum(total_claims, na.rm = TRUE),
    exposure = sum(total_exposure, na.rm = TRUE),
    profit = premium - incurred,
    loss_ratio = incurred / premium,
    claim_frequency = claims / exposure,
    claim_severity = incurred / claims,
    .groups = "drop"
  )

policy_type_summary <- motor_clean %>%
  group_by(policy_type) %>%
  summarise(
    n = n(),
    premium = sum(total_premium, na.rm = TRUE),
    incurred = sum(total_incurred, na.rm = TRUE),
    claims = sum(total_claims, na.rm = TRUE),
    exposure = sum(total_exposure, na.rm = TRUE),
    profit = premium - incurred,
    loss_ratio = incurred / premium,
    claim_frequency = claims / exposure,
    claim_severity = incurred / claims,
    .groups = "drop"
  ) %>%
  arrange(desc(loss_ratio))

policy_status_summary <- motor_clean %>%
  group_by(policy_status) %>%
  summarise(
    n = n(),
    premium = sum(total_premium, na.rm = TRUE),
    incurred = sum(total_incurred, na.rm = TRUE),
    profit = premium - incurred,
    loss_ratio = incurred / premium,
    .groups = "drop"
  )

payment_summary <- motor_clean %>%
  group_by(payment_frequency) %>%
  summarise(
    n = n(),
    premium = sum(total_premium, na.rm = TRUE),
    incurred = sum(total_incurred, na.rm = TRUE),
    claims = sum(total_claims, na.rm = TRUE),
    exposure = sum(total_exposure, na.rm = TRUE),
    profit = premium - incurred,
    loss_ratio = incurred / premium,
    claim_frequency = claims / exposure,
    claim_severity = incurred / claims,
    .groups = "drop"
  )

fuel_summary <- motor_clean %>%
  group_by(fuel_type) %>%
  summarise(
    n = n(),
    premium = sum(total_premium, na.rm = TRUE),
    incurred = sum(total_incurred, na.rm = TRUE),
    claims = sum(total_claims, na.rm = TRUE),
    exposure = sum(total_exposure, na.rm = TRUE),
    profit = premium - incurred,
    loss_ratio = incurred / premium,
    claim_frequency = claims / exposure,
    claim_severity = incurred / claims,
    .groups = "drop"
  )

risk_heatmap <- motor_clean %>%
  group_by(policy_type, payment_frequency) %>%
  summarise(
    n = n(),
    premium = sum(total_premium, na.rm = TRUE),
    incurred = sum(total_incurred, na.rm = TRUE),
    claims = sum(total_claims, na.rm = TRUE),
    exposure = sum(total_exposure, na.rm = TRUE),
    profit = premium - incurred,
    loss_ratio = incurred / premium,
    claim_frequency = claims / exposure,
    .groups = "drop"
  )
```

***Executive Summary***

**Motor Insurance Portfolio Visual Analytics**

Profitability, claim volatility, and high-risk segments

**Main Objective**

-   Identify whether weak profitability is driven by claim frequency, claim severity, or pricing inadequacy.
-   Locate high-risk policy and vehicle segments.
-   Translate visual analytics findings into pricing and underwriting implications.

## Insight 1
***Insight 1 - Portfolio Profitability Deteriorated in 2024***

```{r}
ggplot(year_summary, aes(x = factor(year), y = loss_ratio)) +
  geom_col(fill = "#2C7FB8") +
  geom_hline(
    yintercept = 1,
    linetype = "dashed",
    linewidth = 0.8,
    color = "white"
  ) +
  geom_text(
    aes(label = percent(loss_ratio, accuracy = 0.1)),
    vjust = -0.3,
    size = 5,
    color = "white"
  ) +
  scale_y_continuous(
    labels = percent,
    limits = c(0, 1.1)
  ) +
  labs(
    title = "Portfolio Loss Ratio by Year",
    subtitle = "The dashed line represents the 100% break-even loss ratio",
    x = "Year",
    y = "Loss Ratio"
  ) +
  theme_minimal(base_size = 18)
```

-   Loss ratio increased from 65.7% in 2022 to 74.7% in 2024.
-   The portfolio remains profitable, but the underwriting margin became thinner.
-   The 2024 deterioration motivates deeper frequency-severity and segment-level analysis.

## Insight 2
***Insight 2 - Premium Growth Did Not Fully Offset Claim Growth***
```{r}
year_long <- year_summary %>%
  select(year, premium, incurred) %>%
  pivot_longer(
    cols = c(premium, incurred),
    names_to = "metric",
    values_to = "amount"
  ) %>%
  mutate(
    metric = recode(
      metric,
      "premium" = "Premium Income",
      "incurred" = "Incurred Claims"
    )
  )

ggplot(year_long, aes(x = factor(year), y = amount, fill = metric)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = dollar) +
  scale_fill_manual(
    values = c(
      "Premium Income" = "#4DBBD5",
      "Incurred Claims" = "#E64B35"
    )
  ) +
  labs(
    title = "Premium Income and Incurred Claims by Year",
    subtitle = "Claim costs increased substantially in 2024",
    x = "Year",
    y = "Amount",
    fill = "Metric"
  ) +
  theme_minimal(base_size = 18)
```

-   Both premium income and incurred claims increased from 2022 to 2024.
-   Incurred claims grew faster relative to premium income in 2024.
-   Business growth does not necessarily improve profitability if claim costs grow faster.

## Insight 3
***Insight 3 - Claim Risk is Concentrated in Large-Loss Policies***
```{r}
claim_concentration <- motor_clean %>%
  filter(total_incurred > 0) %>%
  arrange(desc(total_incurred)) %>%
  mutate(
    claim_rank = row_number(),
    claim_percentile = claim_rank / n(),
    cumulative_incurred = cumsum(total_incurred),
    cumulative_share = cumulative_incurred / sum(total_incurred)
  )

top10_share <- claim_concentration %>%
  filter(claim_percentile <= 0.10) %>%
  summarise(share = max(cumulative_share)) %>%
  pull(share)

ggplot(claim_concentration,
       aes(x = claim_percentile, y = cumulative_share)) +
  geom_line(color = "#4DBBD5", linewidth = 1.2) +
  geom_vline(xintercept = 0.10, linetype = "dashed", color = "white") +
  geom_hline(yintercept = top10_share, linetype = "dashed", color = "white") +
  annotate(
    "text",
    x = 0.36,
    y = top10_share + 0.06,
    label = paste0(
      "Top 10% of claim policies account for ",
      percent(top10_share, accuracy = 0.1),
      " of incurred claims"
    ),
    size = 5,
    color = "white"
  ) +
  scale_x_continuous(labels = percent) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Claim Concentration Curve",
    subtitle = "A small share of high-loss policies contributes a large share of total incurred claims",
    x = "Share of Claim Policies, Ordered by Claim Size",
    y = "Cumulative Share of Total Incurred Claims"
  ) +
  theme_minimal(base_size = 18)
```

-   Claim costs are highly concentrated among high-loss policies.
-   Portfolio volatility is strongly influenced by tail-risk observations.
-   Large-loss policies should be monitored separately from average claim behavior.

## Insight 4a
***Insight 4 - COMP_N is the Highest-Risk Policy Type***

```{r}
ggplot(policy_type_summary,
       aes(
         x = reorder(policy_type, loss_ratio),
         y = loss_ratio,
         fill = policy_type == "COMP_N"
       )) +
  geom_col() +
  geom_hline(
    yintercept = 1,
    linetype = "dashed",
    linewidth = 0.8,
    color = "white"
  ) +
  geom_text(
    aes(label = percent(loss_ratio, accuracy = 0.1)),
    hjust = -0.1,
    size = 4.5,
    color = "white"
  ) +
  coord_flip() +
  scale_fill_manual(
    values = c("TRUE" = "#E64B35", "FALSE" = "#4DBBD5"),
    guide = "none"
  ) +
  scale_y_continuous(
    labels = percent,
    limits = c(0, 1.1)
  ) +
  labs(
    title = "Loss Ratio",
    x = "Policy Type",
    y = NULL
  ) +
  theme_minimal(base_size = 16)
```

## Insight 4b
```{r}
ggplot(policy_type_summary,
       aes(
         x = reorder(policy_type, claim_frequency),
         y = claim_frequency,
         fill = policy_type == "COMP_N"
       )) +
  geom_col() +
  geom_text(
    aes(label = round(claim_frequency, 2)),
    hjust = -0.1,
    size = 4.5,
    color = "white"
  ) +
  coord_flip() +
  scale_fill_manual(
    values = c("TRUE" = "#E64B35", "FALSE" = "#4DBBD5"),
    guide = "none"
  ) +
  scale_y_continuous(
    limits = c(0, max(policy_type_summary$claim_frequency) * 1.15)
  ) +
  labs(
    title = "Claim Frequency",
    x = "Policy Type",
    y = NULL
  ) +
  theme_minimal(base_size = 16)
```

-   COMP_N has the highest loss ratio at 93.5%.
-   It also has the highest claim frequency at 1.04.
-   This suggests that COMP_N is a frequency-driven high-risk segment.

## Insight 5
***Insight 5 - Policy Status C is Loss-Making***
```{r}
ggplot(policy_status_summary,
       aes(
         x = policy_status,
         y = loss_ratio,
         fill = loss_ratio > 1
       )) +
  geom_col(width = 0.65) +
  geom_hline(
    yintercept = 1,
    linetype = "dashed",
    linewidth = 0.8,
    color = "white"
  ) +
  geom_text(
    aes(label = percent(loss_ratio, accuracy = 0.1)),
    vjust = -0.3,
    size = 5,
    color = "white"
  ) +
  scale_fill_manual(
    values = c("TRUE" = "#E64B35", "FALSE" = "#4DBBD5"),
    guide = "none"
  ) +
  scale_y_continuous(
    labels = percent,
    limits = c(0, max(policy_status_summary$loss_ratio) * 1.15)
  ) +
  labs(
    title = "Loss Ratio by Policy Status",
    subtitle = "The dashed line represents the 100% break-even loss ratio",
    x = "Policy Status",
    y = "Loss Ratio"
  ) +
  theme_minimal(base_size = 18)
```

-   Policy status C has a loss ratio of 128.0%.
-   This segment is loss-making because claim costs exceed premium income.
-   Policy closure or cancellation may be associated with adverse claim experience.

## Insight 6
***Insight 6 - Diesel Vehicles are Frequency-Driven Risk***
```{r}
fuel_long <- fuel_summary %>%
  select(fuel_type, loss_ratio, claim_frequency, claim_severity) %>%
  pivot_longer(
    cols = c(loss_ratio, claim_frequency, claim_severity),
    names_to = "metric",
    values_to = "value"
  ) %>%
  mutate(
    metric = recode(
      metric,
      "claim_frequency" = "Claim Frequency",
      "claim_severity" = "Claim Severity",
      "loss_ratio" = "Loss Ratio"
    ),
    label = case_when(
      metric == "Claim Frequency" ~ as.character(round(value, 3)),
      metric == "Claim Severity" ~ dollar(value, accuracy = 1),
      metric == "Loss Ratio" ~ percent(value, accuracy = 0.1)
    )
  )

ggplot(fuel_long, aes(x = fuel_type, y = value, fill = fuel_type)) +
  geom_col(width = 0.7) +
  geom_text(
    aes(label = label),
    vjust = -0.35,
    size = 4.2,
    color = "white"
  ) +
  facet_wrap(~ metric, scales = "free_y") +
  scale_fill_manual(
    values = c(
      "D" = "#E64B35",
      "G" = "#4DBBD5"
    )
  ) +
  labs(
    title = "Fuel Type and Claim Risk",
    subtitle = "Diesel vehicles show weaker profitability mainly because of higher claim frequency",
    x = "Fuel Type",
    y = NULL
  ) +
  theme_minimal(base_size = 17) +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )
```

-   Diesel vehicles have higher claim frequency than gasoline vehicles.
-   Claim severity is almost identical between the two groups.
-   The fuel-type difference is mainly frequency-driven rather than severity-driven.

## Insight 7
***Insight 7 - High Risk is Concentrated in COMP_N + Semi-Annual Payment***
```{r}
ggplot(risk_heatmap,
       aes(x = payment_frequency, y = policy_type, fill = loss_ratio)) +
  geom_tile(color = "white") +
  geom_text(
    aes(label = percent(loss_ratio, accuracy = 0.1)),
    size = 5,
    color = "white"
  ) +
  scale_fill_gradient(
    low = "#FEE5D9",
    high = "#CB181D",
    labels = percent,
    name = "Loss Ratio"
  ) +
  labs(
    title = "Risk Heatmap by Policy Type and Payment Frequency",
    subtitle = "COMP_N with semi-annual payment is loss-making",
    x = "Payment Frequency",
    y = "Policy Type"
  ) +
  theme_minimal(base_size = 18)
```

-   COMP_N + S has a loss ratio of 108.8%.
-   This specific product-payment combination is loss-making.
-   Segment-level visual analytics reveals risks hidden in one-variable summaries.

## Insight 8
***Insight 8 - Premium Adequacy Quadrant Confirms COMP_N Risk***
```{r}
ggplot(policy_type_summary,
       aes(
         x = claim_frequency,
         y = loss_ratio,
         size = premium,
         label = policy_type
       )) +
  geom_point(
    alpha = 0.8,
    color = "#4DBBD5"
  ) +
  geom_text(
    vjust = -1,
    size = 5,
    color = "white"
  ) +
  geom_hline(
    yintercept = portfolio_summary$overall_loss_ratio,
    linetype = "dashed",
    linewidth = 0.8,
    color = "white"
  ) +
  geom_vline(
    xintercept = portfolio_summary$overall_claim_frequency,
    linetype = "dashed",
    linewidth = 0.8,
    color = "white"
  ) +
  scale_y_continuous(
    labels = percent,
    limits = c(0.50, 1.05)
  ) +
  scale_x_continuous(
    limits = c(0, max(policy_type_summary$claim_frequency) * 1.15)
  ) +
  scale_size_continuous(labels = dollar) +
  labs(
    title = "Premium Adequacy Quadrant by Policy Type",
    subtitle = "Loss ratio versus claim frequency, with bubble size representing premium volume",
    x = "Claim Frequency",
    y = "Loss Ratio",
    size = "Premium"
  ) +
  theme_minimal(base_size = 18)
```

-   COMP_N is located in the high-frequency and high-loss-ratio area.
-   COMP_E is also above the portfolio average on both dimensions.
-   CC and TP appear to be stronger-performing segments.

## Recommendations
***Recommendations***
**Pricing and Underwriting Priorities**

-   Review pricing and underwriting rules for COMP_N.
-   Prioritise investigation of COMP_N + S, since it is loss-making.
-   Monitor policy status C, which has a loss ratio above 100%.
-   Treat diesel vehicles as a frequency-driven risk segment.
-   Add tail-risk monitoring for large-loss policies.
-   Monitor yearly loss ratio movement to detect pricing inadequacy earlier.

## Final Takeaway
***Final Takeaway***
**Visual analytics identifies where profitability problems are concentrated**

-   The portfolio remains profitable overall.
-   Profitability deteriorated in 2024.
-   Claim risk is concentrated in specific segments.
-   COMP_N, policy status C, and COMP_N + S are the most important risk signals.
-   Visual analytics supports more targeted pricing, underwriting, and portfolio monitoring.