---
title: "Take-Home_Ex01"
author: "JUNCHEN LI"
date: "May 18,2026"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: auto
---

# Visual Analytics for Data Discovery: Investigating Profitability, Claim Volatility, and High-Risk Segments

## 1. Introduction

### 1.1 Background

Motor insurance is one of the major business lines in the general insurance industry. For insurers, portfolio profitability depends not only on the amount of premium collected, but also on the frequency and severity of claims. Even when the overall portfolio appears profitable, certain policyholder, vehicle, or policy segments may still generate high losses and create volatility.

In practice, insurers need to identify which customer, vehicle, and policy characteristics are associated with higher claim costs and lower profitability. Visual analytics provides a useful approach because it combines data visualization, statistical summaries, and exploratory analysis to reveal patterns that may not be obvious from aggregate figures alone.

Therefore, this study applies visual analytics techniques to a motor insurance portfolio dataset to investigate profitability, claim volatility, and high-risk segments.

### 1.2 Objective and Analytically Questions

The objective of this take-home exercise is to use appropriate visual analytics methods, tidyverse functions, and ggplot2 extensions to explore the profitability and volatility of a motor insurance portfolio.

Specifically, this report aims to answer the following analytically questions:

-   What are the distributions of premium income, claim cost, claim frequency, and loss ratio in the insurance portfolio?
-   How does portfolio profitability change across years?
-   Is weak profitability mainly driven by claim frequency, claim severity, or pricing inadequacy?
-   Which policy characteristics, such as policy type, policy status, business type, and payment frequency, are associated with higher loss ratios?
-   Which vehicle and driver characteristics, such as fuel type, driver age, vehicle age, and vehicle value, are associated with higher claim risk?
-   Are the observed visual patterns statistically supported by validation methods?

Based on the analysis, this report provides practical insights that may support insurance pricing, underwriting, and portfolio management decisions.

## 2. Getting Started

### 2.1 Setting the Analytically Tools

```{r}
pacman::p_load(
  tidyverse,
  janitor,
  scales,
  ggthemes,
  plotly,
  DT,
  broom
)

theme_set(theme_minimal())
```

The tidyverse family of packages is used for data importing, wrangling, transformation, and visualization. The janitor package is used to standardize column names and check data quality. The scales package is used to format percentage and currency values in charts. The plot package is used to create an interactive visualization as an extension of ggplot2. The broom package is used to present model outputs in a tidy format.

### 2.2 Data Source

The dataset used in this study is a real-world motor vehicle insurance portfolio dataset from a Spanish insurance company specialized in non-life insurance. The dataset contains policy-level information, including policy characteristics, driver characteristics, vehicle characteristics, premium income, claim counts, claim costs, and exposure.

The main variables used in this analysis include total premium, total incurred claims, total claims, total exposure, policy type, policy status, payment frequency, driver age, vehicle age, fuel type, vehicle value, and circulation area.

## 3.Data Wrangling

### 3.1 Importing Data

```{r}
motor_raw <- read_delim(
  "data/Dataset of motor insurance portfolio.csv",
  delim = ";",
  show_col_types = FALSE
)

glimpse(motor_raw)
```

The dataset is imported into the R environment using the read_delim() function because the CSV file uses semicolons as separators. The glimpse() function is used to check the structure, variable types, and sample values of the imported dataset.

### 3.2 Checking Dataset Structure

```{r}
dim(motor_raw)

names(motor_raw)
```

```{r}
motor_raw %>%
  summarise(
    rows = n(),
    columns = ncol(motor_raw),
    unique_insured = n_distinct(insured_id),
    first_year = min(year, na.rm = TRUE),
    last_year = max(year, na.rm = TRUE)
  )
```

The dataset contains policy-level observations across the years 2022, 2023, and 2024. Each observation represents an insurance policy record with information on policy characteristics, vehicle characteristics, driver characteristics, premium income, claim counts, incurred claim costs, and exposure.

### 3.3 Cleaning Column Names

```{r}
motor <- motor_raw %>%
  clean_names()

names(motor)
```

The clean_names() function from the janitor package is applied to standardise the variable names. This step makes the column names easier to reference in later data wrangling and visualisation steps.

### 3.4 Checking Duplicates and Missing Values

```{r}
sum(duplicated(motor))
```

```{r}
missing_summary <- motor %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "missing_count"
  ) %>%
  filter(missing_count > 0) %>%
  arrange(desc(missing_count))

missing_summary
```

A duplicate check is conducted to ensure that the dataset does not contain repeated records. Missing values are also examined for all variables. The main variables with missing values are fuel_type, vehicle_value, vehicle_age, and age_driving_licence. Since the number of missing values is small relative to the total dataset size, these observations are removed from the analytical dataset.

Handing Missing Value:

```{r}
motor_clean <- motor %>%
  drop_na(
    fuel_type,
    vehicle_value,
    vehicle_age,
    age_driving_licence
  )

dim(motor_clean)
```

### 3.5 Creating Analytically Insurance Metrics

```{r}
motor_clean <- motor_clean %>%
  mutate(
    claim_indicator = if_else(total_claims > 0, 1, 0),
    profit = total_premium - total_incurred,
    loss_ratio = if_else(total_premium > 0, total_incurred / total_premium, NA_real_),
    claim_frequency = if_else(total_exposure > 0, total_claims / total_exposure, NA_real_),
    claim_severity = if_else(total_claims > 0, total_incurred / total_claims, NA_real_)
  )
```

Several analytically variables are created to support the insurance portfolio analysis. Profit is calculated as total premium minus total incurred claims. Loss ratio is calculated as total incurred claims divided by total premium and is used as the main profitability measure. A higher loss ratio indicates weaker profitability.

Claim frequency is calculated as total claims divided by total exposure, while claim severity is calculated as total incurred claims divided by total claims. These two variables help distinguish whether poor portfolio performance is driven by frequent claims or by large claim amounts.

### 3.6 Grouping Driver Age, Vehicle Age, and Vehicle Value

```{r}
motor_clean <- motor_clean %>%
  mutate(
    driver_age_group = cut(
      driver_age,
      breaks = c(17, 25, 35, 45, 55, 65, Inf),
      labels = c("18-25", "26-35", "36-45", "46-55", "56-65", "66+"),
      right = TRUE
    ),
    
    vehicle_age_group = cut(
      vehicle_age,
      breaks = c(-1, 5, 10, 20, 30, Inf),
      labels = c("0-5", "6-10", "11-20", "21-30", "31+"),
      right = TRUE
    ),
    
    vehicle_value_group = cut(
      vehicle_value,
      breaks = quantile(vehicle_value, probs = seq(0, 1, 0.2), na.rm = TRUE),
      include.lowest = TRUE,
      labels = c("Very Low", "Low", "Medium", "High", "Very High")
    )
  )
```

Driver age, vehicle age, and vehicle value are grouped into categorical variables. This makes the visual analysis easier to interpret and allows the report to identify segment-level risk patterns more clearly.

## 4. Portfolio-Level Visual Analytics

### 4.1 Overall Portfolio Profitability

```{r}
portfolio_summary <- motor_clean %>%
  summarise(
    number_of_records = n(),
    unique_insured = n_distinct(insured_id),
    total_premium = sum(total_premium, na.rm = TRUE),
    total_incurred = sum(total_incurred, na.rm = TRUE),
    total_claims = sum(total_claims, na.rm = TRUE),
    total_exposure = sum(total_exposure, na.rm = TRUE),
    overall_profit = total_premium - total_incurred,
    overall_loss_ratio = total_incurred / total_premium,
    overall_claim_frequency = total_claims / total_exposure,
    overall_claim_severity = total_incurred / total_claims
  )

portfolio_summary
```

The overall portfolio summary provides a baseline for the subsequent analysis. The aggregate loss ratio measures whether the portfolio is profitable overall, while claim frequency and claim severity help explain the source of insurance risk.

### 4.2 Yearly Trend in Loss Ratio

```{r}
year_summary <- motor_clean %>%
  group_by(year) %>%
  summarise(
    premium = sum(total_premium, na.rm = TRUE),
    incurred = sum(total_incurred, na.rm = TRUE),
    claims = sum(total_claims, na.rm = TRUE),
    exposure = sum(total_exposure, na.rm = TRUE),
    profit = premium - incurred,
    loss_ratio = incurred / premium,
    claim_frequency = claims / exposure,
    claim_severity = incurred / claims,
    .groups = "drop"
  )

year_summary
```

::: panel-tabset
### The plot

```{r, echo=FALSE}
ggplot(year_summary, aes(x = factor(year), y = loss_ratio)) +
  geom_col() +
  geom_text(
    aes(label = percent(loss_ratio, accuracy = 0.1)),
    vjust = -0.3,
    size = 4
  ) +
  scale_y_continuous(labels = percent, limits = c(0, max(year_summary$loss_ratio) * 1.15)) +
  labs(
    title = "Portfolio Loss Ratio by Year",
    subtitle = "Loss ratio increased over time, suggesting weaker profitability",
    x = "Year",
    y = "Loss Ratio"
  )
```

### The Code

```{r, eval=FALSE}
ggplot(year_summary, aes(x = factor(year), y = loss_ratio)) +
  geom_col() +
  geom_text(
    aes(label = percent(loss_ratio, accuracy = 0.1)),
    vjust = -0.3,
    size = 4
  ) +
  scale_y_continuous(labels = percent, limits = c(0, max(year_summary$loss_ratio) * 1.15)) +
  labs(
    title = "Portfolio Loss Ratio by Year",
    subtitle = "Loss ratio increased over time, suggesting weaker profitability",
    x = "Year",
    y = "Loss Ratio"
  )
```
:::

The yearly loss ratio shows how portfolio profitability changes over time. An increasing loss ratio indicates that claim costs are growing faster than premium income, which may suggest weaker pricing adequacy or higher claim risk.

### 4.3 Premium Growth versus Claim Cost Growth

::: panel-tabset
#### The plot
```{r, echo=FALSE}
year_long <- year_summary %>%
  select(year, premium, incurred) %>%
  pivot_longer(
    cols = c(premium, incurred),
    names_to = "metric",
    values_to = "amount"
  )

ggplot(year_long, aes(x = factor(year), y = amount, fill = metric)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = dollar) +
  labs(
    title = "Premium Income and Incurred Claims by Year",
    subtitle = "Comparing premium income with claim cost over time",
    x = "Year",
    y = "Amount",
    fill = "Metric"
  )
```


### The Code
```{r, eval=FALSE}
year_long <- year_summary %>%
  select(year, premium, incurred) %>%
  pivot_longer(
    cols = c(premium, incurred),
    names_to = "metric",
    values_to = "amount"
  )

ggplot(year_long, aes(x = factor(year), y = amount, fill = metric)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = dollar) +
  labs(
    title = "Premium Income and Incurred Claims by Year",
    subtitle = "Comparing premium income with claim cost over time",
    x = "Year",
    y = "Amount",
    fill = "Metric"
  )
```
:::

This chart compares total premium income and total incurred claims across years. If incurred claims increase faster than premium income, the portfolio loss ratio will rise and profitability will deteriorate.

## 5. Claim Distribution and Volatility

### 5.1 Distribution of Total Incurred Claims

::: panel-tabset
### The plot
```{r, echo=FALSE}
ggplot(
  motor_clean %>% filter(total_incurred > 0),
  aes(x = total_incurred)
) +
  geom_histogram(bins = 50) +
  scale_x_log10(labels = dollar) +
  labs(
    title = "Distribution of Total Incurred Claims",
    subtitle = "Claim costs are shown on a log scale because the distribution is highly skewed",
    x = "Total Incurred Claims, Log Scale",
    y = "Number of Policies"
  )
```

### The Code
```{r, eval=FALSE}
ggplot(
  motor_clean %>% filter(total_incurred > 0),
  aes(x = total_incurred)
) +
  geom_histogram(bins = 50) +
  scale_x_log10(labels = dollar) +
  labs(
    title = "Distribution of Total Incurred Claims",
    subtitle = "Claim costs are shown on a log scale because the distribution is highly skewed",
    x = "Total Incurred Claims, Log Scale",
    y = "Number of Policies"
  )
```
:::

The distribution of total incurred claims is highly right-skewed. Most policies have relatively small claim costs, while a small number of policies generate very large losses. This suggests that portfolio volatility may be driven by tail-risk observations.

### 5.2 Distribution of Loss Ratio

::: panel-tabset
### The plot
```{r, echo=FALSE}
loss_ratio_cap <- quantile(motor_clean$loss_ratio, 0.99, na.rm = TRUE)

ggplot(
  motor_clean %>% filter(loss_ratio >= 0, loss_ratio <= loss_ratio_cap),
  aes(x = loss_ratio)
) +
  geom_histogram(bins = 50) +
  scale_x_continuous(labels = percent) +
  labs(
    title = "Distribution of Loss Ratio",
    subtitle = "The top 1% of loss ratios are excluded for readability",
    x = "Loss Ratio",
    y = "Number of Policies"
  )
```

### The Code
```{r, eval=FALSE}
loss_ratio_cap <- quantile(motor_clean$loss_ratio, 0.99, na.rm = TRUE)

ggplot(
  motor_clean %>% filter(loss_ratio >= 0, loss_ratio <= loss_ratio_cap),
  aes(x = loss_ratio)
) +
  geom_histogram(bins = 50) +
  scale_x_continuous(labels = percent) +
  labs(
    title = "Distribution of Loss Ratio",
    subtitle = "The top 1% of loss ratios are excluded for readability",
    x = "Loss Ratio",
    y = "Number of Policies"
  )
```
:::

The loss ratio distribution shows that many policies have low or zero claim costs, while a smaller group has very high loss ratios. This confirms that risk is unevenly distributed across the portfolio.

### 5.3 Claim Concentration and Tail Risk

```{r}
claim_concentration <- motor_clean %>%
  filter(total_incurred > 0) %>%
  arrange(desc(total_incurred)) %>%
  mutate(
    claim_rank = row_number(),
    claim_percentile = claim_rank / n(),
    cumulative_incurred = cumsum(total_incurred),
    cumulative_share = cumulative_incurred / sum(total_incurred)
  )

claim_concentration_summary <- claim_concentration %>%
  summarise(
    top_1_percent_share = sum(total_incurred[claim_percentile <= 0.01]) / sum(total_incurred),
    top_5_percent_share = sum(total_incurred[claim_percentile <= 0.05]) / sum(total_incurred),
    top_10_percent_share = sum(total_incurred[claim_percentile <= 0.10]) / sum(total_incurred)
  )

claim_concentration_summary
```

::: panel-tabset
### The plot
```{r, echo=FALSE}
ggplot(claim_concentration,
       aes(x = claim_percentile, y = cumulative_share)) +
  geom_line(linewidth = 1) +
  scale_x_continuous(labels = percent) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Claim Concentration Curve",
    subtitle = "Cumulative share of total incurred claims by largest claim policies",
    x = "Share of Claim Policies, Ordered by Claim Size",
    y = "Cumulative Share of Total Incurred Claims"
  )
```

### The Code
```{r, eval=FALSE}
ggplot(claim_concentration,
       aes(x = claim_percentile, y = cumulative_share)) +
  geom_line(linewidth = 1) +
  scale_x_continuous(labels = percent) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Claim Concentration Curve",
    subtitle = "Cumulative share of total incurred claims by largest claim policies",
    x = "Share of Claim Policies, Ordered by Claim Size",
    y = "Cumulative Share of Total Incurred Claims"
  )
```
:::

The claim concentration analysis measures how much of total incurred claims is explained by the largest claim policies. If a small percentage of policies accounts for a large share of total claims, portfolio volatility is strongly affected by tail-risk observations.


### 5.4 Frequency-Severity Decomposition by Year
```{r}
  
```

